### 2021-04-27 paper/paper.tex
Segunda entrega

\documentclass[12pt]{article}
\usepackage{float}
\usepackage{booktabs}
\usepackage{tabulary}
\usepackage{svg}
\usepackage{tikz}
\usepackage{pgfplots}
\usepackage{indentfirst} %alinhar 1 parágrafo

\usepackage{sbc-template}
\usepackage{graphicx}
\usepackage{dirtytalk}

\usepackage[T1]{fontenc}
\usepackage{mathtools}
\usepackage{blkarray, bigstrut}
\usepackage{gauss}
\usepackage{amsmath}
\usepackage{afterpage}
\usepackage{graphicx,url}

%\usepackage[brazil]{babel}
\usepackage[utf8]{inputenc}

\pgfplotsset{compat=1.10}
\usetikzlibrary{%
    intersections,%
    arrows,%
    decorations.pathmorphing,%
    backgrounds,%
    positioning,%
    fit,%
    petri,%
    calc,%
    through,%
    graphs,%
    shapes.misc,%
    trees,%
    mindmap,%
    shadows,%
    calendar%
}

\sloppy

\title{Assessment of climate change impacts on hydro resources in Três Marias reservoir using machine learning models}

\author{%
    Bruno T. R. V. Silva\inst{1}, %
    Carolina S. Ansélmo\inst{1}, \\%
    Paula C. S. Borba\inst{1}, %
    Wallace S. S. Souza\inst{1}%
}

\address{%
    Department of Civil Engineering -- Aeronautics Institute of Technology\\
    São José dos Campos, Brazil
    \email{bruno.valeriano@ga.ita.br, carolina.anselmo@ga.ita.br}%
    \vspace{-2ex}
    \email{paula.borba@ga.ita.br,  wallace.souza@ga.ita.br}
%	\today
}

\begin{document}

\maketitle

%\begin{resumo}

%\end{resumo}

\section{Introduction}

Impacts of climate change on hydro resources have become a growing concern, since it may affect negatively the electricity supply in coming years, particularly in countries highly dependent on hydropower, like Brazil. Therefore, the ability to assess and make predictions about the water level and streamflow of reservoirs allows important decisions to be made by competent authorities in order to ensure that water resources can be appropriately allocated.

% TODO: preencher citação, transformar 'ONS (2020)' em citação
Over 63\% of the whole electricity supply in Brazil came from hydropower in 2019. However, Brazil faces reduced availability of hydro resources in recent years. ONS affirmed that Brazil is experiencing a drought period similar to the period between 1948 to 1955 \cite{onsseca}, while other studies suggested the occurrence of more frequent and anomalous droughts in Brazil due to climate changes.

% TODO: No final do paragrafo, talvez seja bom elencar os estudos que sugerem uma frequência anormal
Climate change may affect hydroelectric production mainly because of the precipitation altering. Reduced precipitation may compromise river flows, while increased precipitation could render a power plant unproductive as it could exceed its capacity. Also, the increase in average temperature may alter the soil moisture level, interfering with the flow and storage of water in dams \cite{Mukheibir2013}. According to \cite{Lyra2018}, southeastern Brazil will suffer a significant reduction of precipitation during the twenty-first century. While \cite{Queiroz2016} noted that the Southeast may be benefited from climate changes with considerable increases in electricity generation, but in other regions of Brazil, climate change will likely cause the reduction of hydroelectricity production.

Global Climate Models (GCMs) simulate interactions of climate systems based on physical principles and observational data. They are tools that are used to predict weather patterns and to explore climate sensitivities in scenarios of climate change \cite{gcm}. To provide long-term simulations of climate change at higher resolution, GCMs are downscaled to Regional Climate Models (RCMs), which include projections at a higher resolution.
This study explored the impact of two climate change scenarios,  Representative Concentration Pathway (RCP 4.5 and RCP 8.5), on hydro resources in the Três Marias reservoir, located in the Southeast of Brazil.

Thus, the objective of this article is to make a forecast of the flow and water level of a reservoir using machine learning predictive models. The dataset come from a simulation of the Platform Project of the Center for Weather Forecasting and Climate Studies (CPTEC) of the National Institute for Space Research (INPE). The simulated input data using will be (i) precipitation, (ii) evaporation, (iii) temperature, and (iv) surface runoff. As output data, the water level and stream flow of the reservoir.

\section{Related bibliography}

As pointed by \cite{ardabili2019deep}, using physical or statistical models to study hydrological processes can raise several problems, such as high computational costs, weakness in uncertainty analysis and the need for a large amount of comprehensive data. Therefore, the usage of machine learning methods have been growing among research community, as shown in Figure~\ref{fig:num_publications}, as a strategy to tackle some of these problems.
In addition, several researchers have used machine learning to predict hydrological patterns, such as to predict flow in future hydrographic basins considering climate changes \cite{bhatta2019evaluation}, daily flow to river \cite{kassem2020predicting}, and flow of the hydroelectric plant reservoir \cite{essenfelder2020smart}. Table~\ref{tab:variables-literature} shows the variables used in these studies.

\begin{figure}[htb]
    \centering
    \begin{tikzpicture}
        \begin{axis}[
                width=13cm,
                height=4.5cm,
                style={/pgf/number format/set thousands separator={}},
                axis x line=bottom,
                axis y line=left,
                axis line style={shorten <=-10},
                xmin=2009,
                xmax=2022,
                xtick={2010,...,2021},
                ymin=0,
                ymax=120,
                ytick={10, 30,...,110},
                xlabel={Year},
                xticklabel style={rotate=45},
                ylabel={Number of publications}
            ]
            \addplot[black,ybar,fill, fill opacity=0.0, bar width = 0.8] table {
                2021    34
                2020    111
                2019    75
                2018    33
                2017    16
                2016    16
                2015    5
                2014    7
                2013    6
                2012    7
                2011    3
                2010    3
            };
        \end{axis}
    \end{tikzpicture}
    \caption{Number of publications about machine learning applied in the study of hydrological processes. The search string used was ``\texttt{(Machine learning) AND (hydrological))}'' in the Engineering Village database}
    \label{fig:num_publications}
\end{figure}

\begin{table}[htbp]
\centering
\caption{Variables in the related bibliograpy}
\label{tab:variables-literature}
    \begin{tabulary}{\textwidth}{lCCC}
        \toprule
           & Kassem et al (2019) & Bhatta et al (2019) & Essenfelder et al (2020) \\
        \midrule
            Precipitation & X & X & X \\
            Temperature & X & X & X \\
            Relative humidity & X &  & \\
            Wind speed  & X &  & \\
            Landuse classes  &  & X & \\
            Stream flow  & X & X & X \\
        \bottomrule
    \end{tabulary}
\end{table}

\cite{kicsi2007streamflow} presented a comparison between four different neural networks algorithms applied in short term daily stream flow forecast. They concluded that the Levenberg–Marquardt algorithm, when compared with backpro pagation, conjugate gradient and cascade correlation takes a relatively small time for training, while giving the best stream flow forecast, even using correlation analysis as the method to determine the input vector. Concurrently, \cite{petty2018streamflow} assessed the capacity of employing machine learning techniques and big data to impute streamflow using historical and real-time data. Their results demonstrated that applying Streamflow Hydrology Estimate using Machine Mearning (SHEM) models could be used to estimate streamflow and even to help fill missing data from historical datasets.

\section{Material and methods}
Using two different scenarios, RCP 4.5 and RCP 8.5, we explored the long-term effect of climate variables on flow and water level variables. The predictor and outcome variables are described in Figure~\ref{fig:approach}.

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.5\linewidth]{Figures/approach.png}
  \caption{Methodology}
  \label{fig:approach}
\end{figure}

\subsection{Descriptive analysis}
We chose Três Marias reservoir for two main reasons. First because of its importance to the Brazilian electricity system, since this reservoir holds a hydropower plant of 396 MW of installed capacity. Second, there are available climate data at a high spatial resolution (5 km) for the Southeastern region, where Três Marias is located. Três Marias reservoir is inside the Alto São Francisco sub-basin in Minas Gerais state (Figure \ref{fig:studyarea}).

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.6\linewidth]{Figures/mapa_v0.png}
  \caption{Alto São Franscisco sub-basin.}
  \label{fig:studyarea}
\end{figure}

In order to reduce the complexity, we opted to use the contribution of each river and sub-basin area to the stream flow and water level from \cite{fundep}. Thus, we made an manual classification from all coordinates to 11 points based in the distance to the nearest river which represents the area of influence (Figure \ref{fig:coord}).

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.6\linewidth]{Figures/coord_contribuicao.png}
  \caption{Coordinate points and their contribution rate.}
  \label{fig:coord}
\end{figure}

\subsection{Data}
In order to apply machine learning to forecast streamflow and water level for Três Marias reservoir, we used data generated by CPTEC/INPE and made available on the Projeta Platform \cite{chou2014assessment,chou2014evaluation,Lyra2018}. This data is split into historic observations, ranging from January, 1961 to December, 2005 and projections modeled according to RCP 4.5 and RCP 8.5, ranging from January, 2006 to December, 2099. For our study, since observations of flow and water levels for Três Marias reservoir made available by ONS \cite{onsnivel,onsvazao} start in 1999, historical data collected from the Projeta platform were filtered to start at the same period.

A list of potential predictors is shown in Table~\ref{tab:predictors}. The study area, Alto São Francisco sub-basin, was subdivided in 1672 equally sized regions, and for each of these regions, data regarding temperature, precipitation, evaporation and surface runoff were gathered, according to temporal resolution, historical data and modeled projections shown in Table~\ref{tab:predictors}, which lists the potential predictors alongside to their respective measurement units. Additionally, Table~\ref{tab:outcomes} shows the measurement units, temporal resolution and date range for data regarding streamflow and water level for the Três Marias reservoir.

\begin{table}[htbp]
\centering
\caption{Relation of predictors considered in this study obtained on the Projeta Platform \cite{chou2014assessment,chou2014evaluation,Lyra2018}}
\label{tab:predictors}
    \resizebox{0.8\columnwidth}{!}{%
    \begin{tabulary}{\textwidth}{lcCcCr}
        \toprule
            Variable       & Unit & Temporal resolution & Historical data & Modeled Projections \\
        \midrule
            Temperature (T)    & ºC   & Daily               & 1999-2005       & 2006-2099 \\
            Precipitation (P) & mm   & Daily               & 1999-2005       & 2006-2099 \\
            Evaporation (E)    & mm   & Daily               & 1999-2005       & 2006-2099 \\
            Surface runoff (R) & mm   & Daily               & 1999-2005       & 2006-2099 \\
        \bottomrule
    \end{tabulary}
    }
\end{table}

\begin{table}[htbp]
    \centering
    \caption{Outcome variables made available by ONS \cite{onsnivel,onsvazao}}
    \label{tab:outcomes}
    \resizebox{0.6\columnwidth}{!}{%
    \begin{tabular}{lccr}
        \toprule
            Variable       & Unit & Temporal resolution & Historical data \\
        \midrule
            Streamflow (SF)           & m³/s & Daily               & 1999-2021 \\
            Water level (WL)    & m    & Daily               & 1999-2021 \\
        \bottomrule
    \end{tabular}
    }
    \label{tab:my_label}
\end{table}

%Descrever os dados: números de exemplos, de atributos, os tipos e escalas dos atributos, se é classificação/regressão, etc.. Se for classificação, dizer a proporção de exemplos na classe com mais observações (majoritária).

%\subsection{Técnicas}

%\section{Results}\label{sec:figs}

% Precisamos citar: Integração de dados (merges), Eliminação de atributos e como tratamos valores fora da faixa (RNOF) e limpeza de dados (verificar)

\section{Data preprocessing}

In order to organize the data into attribute-value format, we followed three main steps: data reduction, data integration and data transformation. Originally, we had a large dataset with the variables P, E, T and R related to 1673 coordinate points inside the sub-basin limits in the period between 1999 to 2005.

We aggregated the coordinate points into 11 groups of rivers according to their contribution area to the Três Marias reservoir and removed information that are not necessary to the model, such as the latitude and longitude (Figure \ref{fig:step1}).

Then, in order to reduce the original dataset and transform it into the information that represents all points in their respective group, we considered the daily sum for P and E, and the daily mean for T and R values (Figure \ref{fig:step2}). Thus, we created weather variables of P, E, T and R for 11 groups of rivers instead of 1673 points.

Although time is an important feature relating to weather variability, we did not consider the date as an attribute, since the seasonal changes are already incorporated in the magnitude of the other variables.

\begin{figure}[htbp]
  \centering
  \includegraphics[width=1\linewidth, trim=0cm 0 0 2.2cm,clip=true]{Figures/STEP1.png}
  \caption{Step 1: data reduction.}
  \label{fig:step1}
\end{figure}

\begin{figure}[htbp]
  \centering
  \includegraphics[width=1\linewidth, trim=0cm 0 0 1.5cm,clip=true]{Figures/STEP2.png}
  \caption{Step 2: feature aggregation.}
  \label{fig:step2}
\end{figure}

In data integration, we combined three different databases. Here, we joined weather variables (predictors) with Water Level (WL) and Stream-Flow (SF), which are both features related to the Três Marias reservoir (Figure \ref{fig:step3}). Besides that, they are the target variables. The mathematical representation of merging databases is illustrated in Figure \ref{fig:matriz1}. Finally, we obtained the data matrix in attribute-value format, which is represented in Figure \ref{fig:matriz2}.

\begin{figure}[htbp]
  \centering
  \includegraphics[width=1\linewidth, trim=0cm 0 0 1.2cm,clip=true]{Figures/STEP3.png}
  \caption{Step 3: data integration.}
  \label{fig:step3}
\end{figure}

\begin{figure}[htbp]
  \centering
  \input{Data/matrix1}
  %\includegraphics[width=1\linewidth]{Figures/matriz1.png}
  \caption{Matrix of different databases}
  \label{fig:matriz1}
\end{figure}

\begin{figure}[htbp]
  \centering
  \input{Data/matrix2}
  %\includegraphics[width=1\linewidth]{Figures/matriz2.png}
  \caption{Data matrix}
  \label{fig:matriz2}
\end{figure}

\subsection{Normalization}
The results of the earlier procedures indicate there's values in multiple scales in the dataset. Since the prediction model selected in the next steps can be sensitive to scale features, we normalized the data with the min-max normalization with the mathematical equation represented in \ref{eqn:Normalization}:
\begin{equation}
\label{eqn:Normalization}
x' = \frac{x-min(x)}{max(x)-min(x)}
\end{equation}
where ~$x$ is the present value, ~$min(x)$ is the minimum value of the feature and ~$max(x)$ is the maximum value of the feature.

We applied this procedure for every variable in the dataset. While only the scale in the values changed with the normalization procedure, it is easier to identify the variation in mean and standard deviation between variables, as shown in Figure \ref{fig:Normalized} for the inputs in the first river.

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.6\linewidth, trim=0cm 0 0 .7cm,clip=true]{Figures/Normalização.png}
  \caption{Boxplot of inputs from raw data x normalized data}
  \label{fig:Normalized}
\end{figure}

Other preprocessing procedures, such as feature selection or dimensionality reduction, are not applicable since we filtered the data in the first step (download from Projeta Platform). Therefore, all data is composed of numeric values that make other transformation techniques unnecessary.

In the future, we intend to verify if the variation between measures can generate better results for the prediction model than absolute values, such as comparison between reservoir levels and variation between days.

In conclusion, we converted the database in the format attribute-value with normalized values between features. The "CSV" file extension archive in the annex provides further information about the newly generated dataset.

%\subsection{Machine learning models} \label{subsec:pre-processamento}

%Parte a ser entregue até o dia 21/06.

%\section{Discussion}

%Discutir os resultados obtidos. Parte a ser entregue até o dia 21/06.

%\section{Conclusion}

\bibliographystyle{sbc}
\bibliography{sbc-template}

\end{document}

### 2021-04-27 setup.py

@@ -79,25 +79,24 @@ rcp8_5 = [
 
 historic = [
     {'scenario': 16, 'variable': 'PREC'},  # done
-    # {'scenario': 16, 'variable': 'TP2M'},  # done
-    # {'scenario': 16, 'variable': 'EVTP'},  # done
-    # {'scenario': 16, 'variable': 'RNOF'}  # done
+    {'scenario': 16, 'variable': 'TP2M'},  # done
+    {'scenario': 16, 'variable': 'EVTP'},  # done
+    {'scenario': 16, 'variable': 'RNOF'}  # done
 ]
 
 [historic_d.update(historic_common) for historic_d in historic]
 
 downloads = []
-# downloads += rcp4_5
+downloads += rcp4_5
 # downloads += rcp8_5
-downloads += historic
+# downloads += historic
 
 
 def main():
-    remove_db()
-    create_db()
+    # remove_db()
+    # create_db()
 
-    parse_shapefile()
-    # print(get_projeta_data_async(**downloads[0], latitude=-13, longitude=-49))
+    # # parse_shapefile()
     parse_projeta_data_from_downloads_list()
 
     Session.remove()
@@ -119,7 +118,7 @@ def parse_shapefile():
 
 
 def parse_projeta_data_from_downloads_list():
-    coordinates = session.query(models.Coordinate).all()[:1]
+    coordinates = session.query(models.Coordinate).all()
 
     variables = {
         'PREC': 'precipitation',
@@ -128,7 +127,7 @@ def parse_projeta_data_from_downloads_list():
         'RNOF': 'surface_runoff',
     }
     for download in downloads:
-        with concurrent.futures.ThreadPoolExecutor(50) as executor:
+        with concurrent.futures.ThreadPoolExecutor(5) as executor:
             results = executor.map(build_variables_table, coordinates, [download for _ in coordinates])
             coordinates_count = session.query(functions.count(models.Coordinate.id)).scalar()
             print(coordinates_count)
@@ -154,6 +153,7 @@ def parse_projeta_data_from_downloads_list():
                     except ValueError:
                         continue
 
+                    data['value'] = None if data['value'] < -100 else data['value']
                     try:
                         var = th_session.query(models.Variables)\
                             .filter(models.Variables.coordinate == coordinate)\
@@ -161,14 +161,12 @@ def parse_projeta_data_from_downloads_list():
                             .filter(models.Variables.time == time)\
                             .filter(models.Variables.scenario == download['scenario'])\
                             .one()
[…]

### 2021-04-27 source/build_data/data_downloader/projeta.py

@@ -33,7 +33,7 @@ def get_data(url):
     while status_code != 200 and tries > 0:
         try:
 
-            response_obj = requests.get(url, timeout=30)
+            response_obj = requests.get(url, timeout=100)
             status_code = response_obj.status_code
             # print(url, status_code, again)
             response = response_obj.json()
@@ -61,15 +61,14 @@ def build_urls(scenario, frequency, variable, latitude, longitude, periods):
     return urls
 
 
-def slice_period(start_month, start_year, end_month, end_year):
+def slice_period(start_month, start_year, end_month, end_year, month_step=12, year_step=20):
     periods = []
-    month_step = 12
-    for year in range(start_year, end_year + 1):
+    for year in range(start_year, end_year + 1, year_step + 1):
         m1 = start_month if year == start_year else 1
         m2 = end_month if year + 1 == end_year else 13
         for month in range(m1, m2, month_step):
             final_end_month = month + month_step - 1
-            periods.append((month, year, final_end_month, year))
+            periods.append((month, year, final_end_month, min(year + year_step, end_year)))
 
     return periods
 
@@ -91,7 +90,7 @@ def slice_period2(start_month, start_year, end_month, end_year):
 
 
 def get_data_async(urls):
-    with concurrent.futures.ThreadPoolExecutor(1) as executor:
+    with concurrent.futures.ThreadPoolExecutor(5) as executor:
         datasets = executor.map(get_data, urls)
 
         data = []

### 2021-04-27 paper/paper.tex
Remoção dos erros e avisos para a entrega 2

@@
 
-% TODO: preencher citação, transformar 'ONS (2020)' em citação
 Over 63\% of the whole electricity supply in Brazil came from hydropower in 2019.
@@
 \centering
-\caption{Variables in the related bibliograpy}
+\caption{Variables in the related bibliography}
 \label{tab:variables-literature}
@@
 Concurrently, \cite{petty2018streamflow} assessed the capacity of employing machine learning techniques and big data to impute streamflow using historical and real-time data.
-Their results demonstrated that applying Streamflow Hydrology Estimate using Machine Mearning (SHEM) models could be used to estimate streamflow and even to help fill missing data from historical datasets.
+Their results demonstrated that applying Streamflow Hydrology Estimate using Machine Learning (SHEM) models could be used to estimate streamflow and even to help fill missing data from historical datasets.
 

### 2021-06-18 model_search.py
ainda em testes

@@ -6,6 +6,7 @@ import pandas
 from eli5.sklearn import PermutationImportance
 from IPython import get_ipython
 from mlxtend.feature_selection import ColumnSelector
+from scipy.stats.stats import DescribeResult
 from sklearn.compose import (ColumnTransformer, TransformedTargetRegressor,
                              make_column_selector)
 from sklearn.dummy import DummyRegressor
@@ -24,7 +25,14 @@ from sklearn.tree import DecisionTreeRegressor
 from sqlalchemy import select
 from sqlalchemy.sql.expression import true
 
-from source.custom_transfomers.debug_transformer import Debug
+from source.custom_transfomers.dataframe_transformers import (
+    AggregateTransformer,
+    GroupByTransformer,
+    PivotTransformer,
+    ResetIndexTransformer,
+    TestTargetColumnSelector,
+    TestTransformer)
+from source.custom_transfomers.debug_transformer import Debug, DebugY
 from source.custom_transfomers.time_window_transformer import \
     TimeWindowTransformer
 from source.data_base.connection import session
@@ -56,7 +64,8 @@ query = select(
     models.Reservoir.streamflow
 ).\
     join(models.Variables.coordinate).\
-    join(models.Reservoir, models.Variables.date == models.Reservoir.date)
+    join(models.Reservoir, models.Variables.date == models.Reservoir.date).\
+    order_by(models.Variables.date)
 
 RawDataFrame = pandas.read_sql(query, session.bind)
 
@@ -126,22 +135,31 @@ previous_cols_loc = [ConsolidatedDataFrame.columns.get_loc((col, '')) for col in
 # %%
 cv = KFold(n_splits=3, random_state=seed, shuffle=True)
 
+
 pipeline_steps = [
-    ('columns', ColumnTransformer([
-        (
-            'normalize',
-            MinMaxScaler(feature_range=(0, 1)),
-            ColumnsLoc(var_cols).get_loc
-        ),
-        (
-            'month',
-            OrdinalEncoder(),
-            ColumnsLoc(date_cols).get_loc
-        )
-    ],
-        remainder='drop',
-    )),
-    ('reg', DummyRegressor())
+    ('test', TestTransformer(3)),
+    # ('test_debug', Debug()),
+    ('target_col', TestTargetColumnSelector('level')),
+    ('debug', Debug()),
[…]

### 2021-06-18 source/custom_transfomers/dataframe_transformers.py

@@ -1,9 +1,11 @@
-from typing import Union
+from typing import Any, Set, Type, Union
 
 import numpy as np
-from pandas import DataFrame
-from sklearn.base import BaseEstimator, TransformerMixin
+from pandas import DataFrame, MultiIndex, Index
+import pandas
+from sklearn.base import BaseEstimator, RegressorMixin, TransformerMixin
 from sklearn.utils.validation import check_array, check_is_fitted
+from source.project_utils.data_manipulation import ColumnsLoc
 
 
 class PivotTransformer(TransformerMixin, BaseEstimator):
@@ -18,16 +20,16 @@ class PivotTransformer(TransformerMixin, BaseEstimator):
         if hasattr(self, 'is_fitted_'):
             del self.is_fitted_
 
-        X = check_array(X, accept_sparse=True, force_all_finite='allow-nan', ensure_2d=False)
-
         self.is_fitted_ = True
+
+        self.transform(y)
 
         return self
 
     def transform(self, X: DataFrame, copy=None) -> DataFrame:
         check_is_fitted(self, 'is_fitted_')
 
-        new_X = X if copy == False else X.copy()
+        new_X = X
 
         new_X.pivot(
             index=self.index,
@@ -40,7 +42,7 @@ class PivotTransformer(TransformerMixin, BaseEstimator):
 
 class ResetIndexTransformer(TransformerMixin, BaseEstimator):
 
-    def __init__(self, level=None, drop=False, inplace=False, col_level=0, col_fill='') -> None:
+    def __init__(self, level=None, drop=False, inplace=True, col_level=0, col_fill='') -> None:
         self.level = level
         self.drop = drop
         self.inplace = inplace
@@ -52,24 +54,26 @@ class ResetIndexTransformer(TransformerMixin, BaseEstimator):
         if hasattr(self, 'is_fitted_'):
             del self.is_fitted_
 
-        X = check_array(X, accept_sparse=True, force_all_finite='allow-nan', ensure_2d=False)
-
         self.is_fitted_ = True
+
+        self.transform(y)
 
         return self
 
     def transform(self, X: DataFrame, copy=None) -> DataFrame:
         check_is_fitted(self, 'is_fitted_')
 
-        new_X = X if copy == False else X.copy()
-
-        new_X.pivot(
-            index=self.index,
-            columns=self.columns,
-            values=self.values
+        temp_X = X.reset_index(
+            level=self.level,
+            drop=self.drop,
+            inplace=self.inplace,
[…]

### 2021-06-18 source/custom_transfomers/debug_transformer.py

@@ -7,8 +7,31 @@ class Debug(BaseEstimator, TransformerMixin):
 
     def transform(self, X):
         print(DataFrame(X))
+        # print(DataFrame(X).info())
         print(X.shape)
         return X
 
     def fit(self, X, y=None, **fit_params):
         return self
+
+    def inverse_transform(self, X):
+        return X
+
+    def predict(self, X):
+        return X
+
+
+class DebugY(BaseEstimator, TransformerMixin):
+
+    def transform(self, X):
+        return X
+
+    def fit(self, X, y=None, **fit_params):
+        print(y)
+        return self
+
+    def inverse_transform(self, X):
+        return X
+
+    def predict(self, X):
+        return X

### 2021-06-18 model_load.py
setup para extrais params

@@ -30,7 +30,7 @@ from source.custom_transfomers.time_window_transformer import \
     TimeWindowTransformer
 from source.data_base.connection import session
 from source.data_base.models import models
-from source.project_utils.constants import targets
+from source.project_utils.constants import targets, targets_models
 from source.project_utils.data_manipulation import generate_aggregation
 
 pandas.set_option('display.max_columns', 51)
@@ -103,14 +103,14 @@ ConsolidatedDataFrame.insert(0, 'month', months)
 # %%
 # Carregar o modelo
 regression_models = {}
-for target in targets:
+for target, _ in targets_models.items():
     regression_models[target] = joblib.load(f'model/{target}.pkl')
 
 # %%
 df = pandas.DataFrame()
-for target in targets:
-    df[target] = ConsolidatedDataFrame[target]
-    df[f'p_{target}'] = regression_models[target].predict(ConsolidatedDataFrame)
+for target, _ in targets_models.items():
+    df[target.split('_')[0]] = ConsolidatedDataFrame[target.split('-')[0]]
+    df[f'p_{target}'] = regression_models[target]['best_estimator'].predict(ConsolidatedDataFrame)
 df
 
 # %%

### 2021-06-18 model_search copy.py

@@ -1,326 +1,0 @@
-# %%
-import eli5
-import joblib
-import numpy
-import pandas
-from eli5.sklearn import PermutationImportance
-from IPython import get_ipython
-from mlxtend.feature_selection import ColumnSelector
-from sklearn.compose import ColumnTransformer, TransformedTargetRegressor, make_column_selector
-from sklearn.dummy import DummyRegressor
-from sklearn.ensemble import RandomForestRegressor, StackingRegressor
-from sklearn.impute import SimpleImputer
-from sklearn.inspection import permutation_importance
-from sklearn.linear_model import Ridge
-
-from sklearn.metrics import (accuracy_score, f1_score, make_scorer,
-                             precision_score, recall_score)
-from sklearn.model_selection import (GridSearchCV, KFold, cross_val_score,
-                                     cross_validate, train_test_split)
-from sklearn.pipeline import Pipeline
-from sklearn.preprocessing import MinMaxScaler, StandardScaler, OrdinalEncoder
-from sklearn.svm import SVR
-from sklearn.tree import DecisionTreeRegressor
-from sqlalchemy import select
-from sqlalchemy.sql.expression import true
-
-from source.custom_transfomers.debug_transformer import Debug
-from source.custom_transfomers.time_window_transformer import TimeWindowTransformer
-from source.data_base.connection import session
-from source.data_base.models import models
-from source.project_utils.data_manipulation import generate_aggregation, get_columns_loc
-
-pandas.set_option('display.max_columns', 51)
-
-# To add a new cell, type '# %%'
-# To add a new markdown cell, type '# %% [markdown]'
-
-# %%
-# Construção do dataframe utilizando buscas no banco de dados sql
-get_ipython().run_line_magic('load_ext', 'autoreload')
-get_ipython().run_line_magic('autoreload', '2')
-
-
-# %%
-query = select(
-    models.Variables.date,
-    models.Variables.precipitation.label('precipitation'),
-    models.Variables.temperature.label('temperature'),
-    models.Variables.evaporation.label('evaporation'),
-    models.Variables.surface_runoff.label('surface_runoff'),
-    models.Coordinate.river_id.label('river'),
-    models.Reservoir.level,
-    models.Reservoir.streamflow
-).\
-    join(models.Variables.coordinate).\
[…]

### 2021-06-18 model_search.py

@@ -85,17 +85,18 @@ ConsolidatedDataFrame = (
     reset_index(['level', 'streamflow'])
 )
 
+shift = 7
 
 ConsolidatedDataFrame.insert(
     loc=ConsolidatedDataFrame.columns.get_loc(('level', '')) + 1,
     column='previous_level',
-    value=pandas.DataFrame(ConsolidatedDataFrame['level']).shift(1).values
+    value=pandas.DataFrame(ConsolidatedDataFrame['level']).shift(shift).values
 )
 
 ConsolidatedDataFrame.insert(
     loc=ConsolidatedDataFrame.columns.get_loc(('streamflow', '')) + 1,
     column='previous_streamflow',
-    value=pandas.DataFrame(ConsolidatedDataFrame['streamflow']).shift(1).values
+    value=pandas.DataFrame(ConsolidatedDataFrame['streamflow']).shift(shift).values
 )
 
 ConsolidatedDataFrame = ConsolidatedDataFrame.dropna()
@@ -137,44 +138,45 @@ cv = KFold(n_splits=3, random_state=seed, shuffle=True)
 
 
 pipeline_steps = [
-    ('test', TestTransformer(3)),
-    # ('test_debug', Debug()),
-    ('target_col', TestTargetColumnSelector('level')),
-    ('debug', Debug()),
-    ('reg', TransformedTargetRegressor(
-        transformer=Debug(),
-        regressor=DecisionTreeRegressor(max_depth=5))
-     )
-    # ('columns', ColumnTransformer([
-    #     (
-    #         'normalize',
-    #         MinMaxScaler(feature_range=(0, 1)),
-    #         ColumnsLoc(var_cols).get_loc
-    #     ),
-    #     (
-    #         'month',
-    #         OrdinalEncoder(),
-    #         ColumnsLoc(date_cols).get_loc
-    #     )
-    # ],
-    #     remainder='drop',
-    # )),
-
+    ('columns', ColumnTransformer([
+        (
+            'normalize',
+            MinMaxScaler(feature_range=(0, 1)),
+            ColumnsLoc(var_cols).get_loc
+        ),
+        (
+            'month',
+            OrdinalEncoder(),
+            ColumnsLoc(date_cols).get_loc
+        ),
+        (
+            'historical',
+            'passthrough',
+            ColumnsLoc(previous_cols).get_loc
+        )
+    ],
+        remainder='drop',
+    )),
+    ('reg', DummyRegressor())
 ]
 
 grid_search_params = dict(
[…]

### 2021-06-18 source/project_utils/constants.py

@@ -1,10 +1,10 @@
 targets = [
-    'level',
-    # 'streamflow',
+    # 'level',
+    'streamflow',
     # 'level_variation',
     # 'streamflow_variation'
 ]
 
 targets_models = {
-    'level': '1MFx5l0_stavsU1X_7NMlrpruG_16xP4h'
+    'streamflow-s_d07': '1ogxLbQdyn4gQu5MXqReHKQC4QuP3SjOz'
 }

### 2021-06-18 source/setup.py

@@ -1,7 +1,7 @@
 from google_drive_downloader import GoogleDriveDownloader as gdd
 from project_utils.constants import targets, targets_models
 
-for target in targets:
+for target, _ in targets_models.items():
     gdd.download_file_from_google_drive(
         file_id=targets_models[target],
         dest_path=f'model/{target}.pkl',

### 2021-06-18 model_search.py
Atualização nos params do random forest para o stacked com final SVR

@@ -208,7 +208,7 @@ grid_search_params = dict(
                 )
             ],
             'reg__RandomForest__random_state': [seed],
-            'reg__RandomForest__n_estimators': [1000],
+            'reg__RandomForest__n_estimators': [10, 100, 1000],
             'reg__SVR__regressor__C': range(15, 31, 5),
             'reg__final_estimator__regressor__C': range(7, 16, 5),
         },

### 2021-06-18 model_search.py
Gridsearch verboso

@@ -233,6 +233,7 @@ grid_search_params = dict(
     scoring='r2',
     cv=cv,
     n_jobs=-1,
+    verbose=10,
     refit=False
 )
 

### 2021-06-18 model_search.py
Kfold para 10 no GridsearchCV

@@ -134,7 +134,7 @@ previous_cols = ['previous_level', 'previous_streamflow']
 previous_cols_loc = [ConsolidatedDataFrame.columns.get_loc((col, '')) for col in previous_cols]
 
 # %%
-cv = KFold(n_splits=3, random_state=seed, shuffle=True)
+cv = KFold(n_splits=10, random_state=seed, shuffle=True)
 
 
 pipeline_steps = [

### 2021-06-18 Setup.sh
Ajuste de alguns parâmetros e melhoria no readme

@@ -1,1 +1,0 @@
-.env/Scripts/python.exe source/setup.py

### 2021-06-18 model_search.py

@@ -85,7 +85,7 @@ ConsolidatedDataFrame = (
     reset_index(['level', 'streamflow'])
 )
 
-shift = 7
+shift = 5
 
 ConsolidatedDataFrame.insert(
     loc=ConsolidatedDataFrame.columns.get_loc(('level', '')) + 1,
@@ -268,7 +268,7 @@ for target in targets:
         best_estimator=reg_search[target].estimator,
         estimators={},
     )
-    for roll in [1]:  # range(1, 22, 10):
+    for roll in range(1, 22, 10):
         WindowedDataFrame = TimeWindowTransformer(var_cols, roll, agg, True).fit_transform(ConsolidatedDataFrame)
         regressor[target]['estimators'][roll] = reg_search[target].fit(WindowedDataFrame, WindowedDataFrame[target])
 

### 2021-06-18 source/setup.py

@@ -1,5 +1,5 @@
 from google_drive_downloader import GoogleDriveDownloader as gdd
-from project_utils.constants import targets, targets_models
+from project_utils.constants import targets_models
 
 for target, _ in targets_models.items():
     gdd.download_file_from_google_drive(

### 2021-06-19 model_search.py
Setup para rodar todos os testes estipulados

@@ -85,18 +85,19 @@ ConsolidatedDataFrame = (
     reset_index(['level', 'streamflow'])
 )
 
-shift = 5
+months = [date.month for date in ConsolidatedDataFrame.index.get_level_values('date')]
+ConsolidatedDataFrame.insert(0, 'month', months)
 
 ConsolidatedDataFrame.insert(
     loc=ConsolidatedDataFrame.columns.get_loc(('level', '')) + 1,
-    column='previous_level',
-    value=pandas.DataFrame(ConsolidatedDataFrame['level']).shift(shift).values
+    column='yesterday_level',
+    value=pandas.DataFrame(ConsolidatedDataFrame['level']).shift(1).values
 )
 
 ConsolidatedDataFrame.insert(
     loc=ConsolidatedDataFrame.columns.get_loc(('streamflow', '')) + 1,
-    column='previous_streamflow',
-    value=pandas.DataFrame(ConsolidatedDataFrame['streamflow']).shift(shift).values
+    column='yesterday_streamflow',
+    value=pandas.DataFrame(ConsolidatedDataFrame['streamflow']).shift(1).values
 )
 
 ConsolidatedDataFrame = ConsolidatedDataFrame.dropna()
@@ -104,17 +105,14 @@ ConsolidatedDataFrame = ConsolidatedDataFrame.dropna()
 ConsolidatedDataFrame.insert(
     loc=ConsolidatedDataFrame.columns.get_loc(('level', '')) + 2,
     column='level_variation',
-    value=ConsolidatedDataFrame.level - ConsolidatedDataFrame.previous_level
+    value=ConsolidatedDataFrame.level - ConsolidatedDataFrame.yesterday_level
 )
 
 ConsolidatedDataFrame.insert(
     loc=ConsolidatedDataFrame.columns.get_loc(('streamflow', '')) + 2,
     column='streamflow_variation',
-    value=ConsolidatedDataFrame.streamflow - ConsolidatedDataFrame.previous_streamflow
+    value=ConsolidatedDataFrame.streamflow - ConsolidatedDataFrame.yesterday_streamflow
 )
-
-months = [date.month for date in ConsolidatedDataFrame.index.get_level_values('date')]
-ConsolidatedDataFrame.insert(0, 'month', months)
 
 
 # %%
@@ -124,14 +122,10 @@ scorer = make_scorer(accuracy_score)
 rivers = session.query(models.River).all()
 
[…]

### 2021-06-19 source/project_utils/constants.py

@@ -1,10 +1,11 @@
 targets = [
-    # 'level',
+    'level',
     'streamflow',
     # 'level_variation',
     # 'streamflow_variation'
 ]
 
 targets_models = {
-    'streamflow-s_d07': '1ogxLbQdyn4gQu5MXqReHKQC4QuP3SjOz'
+    'streamflow-s_d07': '1ogxLbQdyn4gQu5MXqReHKQC4QuP3SjOz',
+    'streamflow-s_d05': '1PH2NuRLeTZKGGk5GXfQb0eN3npXa90r5',
 }

### 2021-06-19 model_load.py
códigos e modelos para shifts e janelamentos de 1, 15 e 30 dias

@@ -106,11 +106,5 @@ regression_models = {}
 for target, _ in targets_models.items():
     regression_models[target] = joblib.load(f'model/{target}.pkl')
 
-# %%
-df = pandas.DataFrame()
-for target, _ in targets_models.items():
-    df[target.split('_')[0]] = ConsolidatedDataFrame[target.split('-')[0]]
-    df[f'p_{target}'] = regression_models[target]['best_estimator'].predict(ConsolidatedDataFrame)
-df
 
 # %%

### 2021-06-19 model_search.py

@@ -46,13 +46,14 @@ pandas.set_option('display.max_columns', 51)
 # To add a new cell, type '# %%'
 # To add a new markdown cell, type '# %% [markdown]'
 
+# # %%
+# get_ipython().run_line_magic('load_ext', 'autoreload')
+# get_ipython().run_line_magic('autoreload', '2')
+
+
 # %%
 # Construção do dataframe utilizando buscas no banco de dados sql
-get_ipython().run_line_magic('load_ext', 'autoreload')
-get_ipython().run_line_magic('autoreload', '2')
-
-
-# %%
+print('SQL query...')
 query = select(
     models.Variables.date,
     models.Variables.precipitation.label('precipitation'),
@@ -157,9 +158,9 @@ pipeline_steps = [
 grid_search_params = dict(
     param_grid=[
         # Para testes rápidos
-        # {
-        #     'reg': [DecisionTreeRegressor(max_depth=5)]
-        # },
+        {
+            'reg': [DecisionTreeRegressor(max_depth=5)]
+        },
         {
             'reg': [
                 TransformedTargetRegressor(
@@ -188,6 +189,23 @@ grid_search_params = dict(
                 'passthrough'
             ],
         },
+        {
+            'reg': [
+                StackingRegressor(
+                    estimators=[
+                        ('RandomForest', RandomForestRegressor()),
+                        ('SVR', TransformedTargetRegressor(
+                            transformer=MinMaxScaler(feature_range=(0, 1)),
+                            regressor=SVR()
+                        ))
+                    ],
+                    final_estimator=Ridge()
+                )
+            ],
+            'reg__RandomForest__random_state': [seed],
+            'reg__RandomForest__n_estimators': [10, 100, 250],
+            'reg__SVR__regressor__C': range(15, 31, 5),
+        },
         # {
         #     'reg': [
         #         StackingRegressor(
@@ -209,23 +227,6 @@ grid_search_params = dict(
         #     'reg__SVR__regressor__C': range(15, 31, 5),
         #     'reg__final_estimator__regressor__C': range(7, 16, 5),
         # },
-        {
-            'reg': [
-                StackingRegressor(
-                    estimators=[
-                        ('RandomForest', RandomForestRegressor()),
-                        ('SVR', TransformedTargetRegressor(
[…]

### 2021-06-19 source/project_utils/constants.py

@@ -6,6 +6,7 @@ targets = [
 ]
 
 targets_models = {
-    'streamflow-s_d07': '1ogxLbQdyn4gQu5MXqReHKQC4QuP3SjOz',
-    'streamflow-s_d05': '1PH2NuRLeTZKGGk5GXfQb0eN3npXa90r5',
+    'streamflow-s_d01': '1cmib3ogVDYTWISmUy3BL77HO_3H2PxLT',
+    'streamflow-s_d015': '1nLVWmqW8Uayrrq5xEX3flwH-bIPaaBJ9',
+    'streamflow-s_d030': '1gKCdINSDXNN9jDMidsjB7Iw4c2JtRNto',
 }

### 2021-06-19 model_load.py
Modelos para mais dias de variações no shift

@@ -13,6 +13,7 @@ from sklearn.compose import ColumnTransformer, TransformedTargetRegressor
 from sklearn.dummy import DummyRegressor
 from sklearn.ensemble import RandomForestRegressor, StackingRegressor
 from sklearn.impute import SimpleImputer
+from sklearn.inspection import permutation_importance
 from sklearn.linear_model import Ridge
 from sklearn.metrics import (accuracy_score, f1_score, make_scorer,
                              precision_score, recall_score)
@@ -106,5 +107,28 @@ regression_models = {}
 for target, _ in targets_models.items():
     regression_models[target] = joblib.load(f'model/{target}.pkl')
 
+# # %%
+# # Testar importância das features
+# for target in targets:
+#     regression_models[target]['permutation_importance'] = permutation_importance(
+#         target_regressor[target]['best_estimator'],
+#         target_regressor[target]['windowed_data'],
+#         target_regressor[target]['windowed_data'][target],
+#         n_repeats=10,
+#         random_state=0,
+#         n_jobs=-1)
+
+# # %%
+# # Printar importância das features
+# for target in targets:
+#     print(pandas.DataFrame(target_regressor[target]['permutation_importance'].importances_mean))
+
+# %%
+for key, item in regression_models.items():
+    for i in [1, 15, 30]:
+        mean, std = pandas.DataFrame(item['estimators'][i].cv_results_).sort_values(
+            'rank_test_r2')[['mean_test_r2', 'std_test_r2']][:1].values[0]
+        print(f'------{key}-------')
+        print(f'{i} -', f'{mean:.05f} +- {std:.05f}')
 
 # %%

### 2021-06-19 model_search.py

@@ -243,7 +243,7 @@ grid_search_params = dict(
     ],
     cv=cv,
     n_jobs=-1,
-    verbose=10,
+    verbose=2,
     error_score="raise",
     refit='r2'
 )
@@ -262,7 +262,7 @@ agg.update(runoff_agg)
 # %%
 print('GridSearch...')
 shift_regressors = {}
-for shift in [1, 15, 30]:
+for shift in [1, 8, 15, 22, 29, 30]:
     ShiftedDataFrame = ConsolidatedDataFrame.copy()
 
     ShiftedDataFrame.insert(
@@ -299,6 +299,6 @@ for shift in [1, 15, 30]:
                 target_regressor[target]['best_windowing'] = roll
                 target_regressor[target]['windowed_data'] = WindowedDataFrame
 
-        joblib.dump(target_regressor[target], f'model/{target}-s_d0{shift}.pkl')
+        joblib.dump(target_regressor[target], f'model/{target}-s_d{shift:02d}.pkl')
 
 # %%

### 2021-06-19 source/project_utils/constants.py

@@ -7,6 +7,9 @@ targets = [
 
 targets_models = {
     'streamflow-s_d01': '1cmib3ogVDYTWISmUy3BL77HO_3H2PxLT',
-    'streamflow-s_d015': '1nLVWmqW8Uayrrq5xEX3flwH-bIPaaBJ9',
-    'streamflow-s_d030': '1gKCdINSDXNN9jDMidsjB7Iw4c2JtRNto',
+    'streamflow-s_d08': '1zX38eY128eAofryxBufRYJm3RddmDmnV',
+    'streamflow-s_d15': '1nLVWmqW8Uayrrq5xEX3flwH-bIPaaBJ9',
+    'streamflow-s_d22': '1MikMxtvQAN_T8GEcRs-9sKBxsCoFYGAm',
+    'streamflow-s_d29': '1wDVAzePrwtGlx9JY0zkowgNHv8cpodr2',
+    'streamflow-s_d30': '1gKCdINSDXNN9jDMidsjB7Iw4c2JtRNto',
 }

### 2021-06-20 model_load.py
matriz de importâncias

@@ -124,11 +124,34 @@ for target, _ in targets_models.items():
 #     print(pandas.DataFrame(target_regressor[target]['permutation_importance'].importances_mean))
 
 # %%
-for key, item in regression_models.items():
-    for i in [1, 15, 30]:
+for i in [1, 15, 30]:
+    for key, item in regression_models.items():
+
         mean, std = pandas.DataFrame(item['estimators'][i].cv_results_).sort_values(
             'rank_test_r2')[['mean_test_r2', 'std_test_r2']][:1].values[0]
         print(f'------{key}-------')
         print(f'{i} -', f'{mean:.05f} +- {std:.05f}')
 
 # %%
+# Testar importância das features
+ImportancesDataFrame = pandas.DataFrame()
+for i in [1, 15, 30]:
+    for target in targets_models:
+        importances = permutation_importance(
+            regression_models[target]['best_estimator'],
+            regression_models[target]['windowed_data'],
+            regression_models[target]['windowed_data'][target.split('-')[0]],
+            n_repeats=10,
+            random_state=0,
+            n_jobs=-1)
+        ImportancesDataFrame[f'{target}_{i}-mean'] = importances.importances_mean
+        ImportancesDataFrame[f'{target}_{i}-std'] = importances.importances_std
+# %%
+# Printar importância das features
+
+ImportancesDataFrame.filter(like='mean', axis='columns')
+
+# %%
+ImportancesDataFrame.set_index(
+    regression_models['streamflow-s_d01']['windowed_data'].columns
+).filter(like='mean', axis='columns').iloc[:9].transpose()

### 2021-06-20 model_load.py
limpeza no código

@@ -36,93 +36,11 @@ from source.project_utils.data_manipulation import generate_aggregation
 
 pandas.set_option('display.max_columns', 51)
 
-
-# %%
-query = select(
-    models.Variables.date,
-    models.Variables.precipitation.label('precipitation'),
-    models.Variables.temperature.label('temperature'),
-    models.Variables.evaporation.label('evaporation'),
-    models.Variables.surface_runoff.label('surface_runoff'),
-    models.Coordinate.river_id.label('river'),
-    models.Reservoir.level,
-    models.Reservoir.streamflow
-).\
-    join(models.Variables.coordinate).\
-    join(models.Reservoir, models.Variables.date == models.Reservoir.date)
-
-RawDataFrame = pandas.read_sql(query, session.bind)
-
-
-# %%
-# DataFrame consolidado porém com os atributos para cada rio posicionados em uma diferente coluna
-ConsolidatedDataFrame = (
-    RawDataFrame.
-    groupby(['date', 'level', 'river', 'streamflow']).
-    agg({
-        'precipitation': 'sum',
-        'evaporation': 'sum',
-        'temperature': 'mean',
-        'surface_runoff': 'mean',
-    }).
-    reset_index().
-    pivot(index=["date", 'level', 'streamflow'], columns="river").
-    reset_index(['level', 'streamflow'])
-)
-
-
-ConsolidatedDataFrame.insert(
-    loc=ConsolidatedDataFrame.columns.get_loc(('level', '')) + 1,
-    column='previous_level',
-    value=pandas.DataFrame(ConsolidatedDataFrame['level']).shift(1).values
-)
-
-ConsolidatedDataFrame.insert(
-    loc=ConsolidatedDataFrame.columns.get_loc(('streamflow', '')) + 1,
-    column='previous_streamflow',
-    value=pandas.DataFrame(ConsolidatedDataFrame['streamflow']).shift(1).values
-)
-
-ConsolidatedDataFrame = ConsolidatedDataFrame.dropna()
-
-ConsolidatedDataFrame.insert(
-    loc=ConsolidatedDataFrame.columns.get_loc(('level', '')) + 2,
-    column='level_variation',
-    value=ConsolidatedDataFrame.level - ConsolidatedDataFrame.previous_level
-)
-
-ConsolidatedDataFrame.insert(
-    loc=ConsolidatedDataFrame.columns.get_loc(('streamflow', '')) + 2,
-    column='streamflow_variation',
[…]

### 2021-06-21 model_search.py
Calculo da importância das variáveis

@@ -262,7 +262,7 @@ agg.update(runoff_agg)
 # %%
 print('GridSearch...')
 shift_regressors = {}
-for shift in [1, 8, 15, 22, 29, 30]:
+for shift in [1, 8, 15, 22, 29]:
     ShiftedDataFrame = ConsolidatedDataFrame.copy()
 
     ShiftedDataFrame.insert(
@@ -302,3 +302,37 @@ for shift in [1, 8, 15, 22, 29, 30]:
         joblib.dump(target_regressor[target], f'model/{target}-s_d{shift:02d}.pkl')
 
 # %%
+regression_models = shift_regressors
+for i in [1, 15, 30]:
+    for score_name, item in regression_models.items():
+
+        mean, std = pandas.DataFrame(item['estimators'][i].cv_results_).sort_values(
+            'rank_test_r2')[['mean_test_r2', 'std_test_r2']][:1].values[0]
+        print(f'------{score_name}-------')
+        print(f'{i} -', f'{mean:.05f} +- {std:.05f}')
+
+# %%
+# Testar importância das features
+ImportancesDataFrame = pandas.DataFrame()
+names = ['target', 'windowing', 'fun']
+columns = regression_models['streamflow-s_d01']['windowed_data'].columns
+for i in [1, 15, 30]:
+    for target in [1, 8, 15, 22, 29]:
+        lagging = int(target[-2:])
+        importances = permutation_importance(
+            regression_models[target]['estimators'][i].best_estimator_,
+            regression_models[target]['windowed_data'],
+            regression_models[target]['windowed_data'][target.split('-')[0]],
+            n_repeats=10,
+            random_state=0,
+            n_jobs=-1)
+
+        mean_index = pandas.MultiIndex.from_tuples([(lagging, i, 'mean')], names=names)
+        std_index = pandas.MultiIndex.from_tuples([(lagging, i, 'std')], names=names)
+        mean_df = pandas.DataFrame(importances.importances_mean, columns=mean_index, index=columns).transpose()
+        std_df = pandas.DataFrame(importances.importances_std, columns=std_index, index=columns).transpose()
+        ImportancesDataFrame = ImportancesDataFrame.append(mean_df)
+        ImportancesDataFrame = ImportancesDataFrame.append(std_df)
+
+# %%
+joblib.dump(ImportancesDataFrame, 'model/importances.pkl')

### 2021-06-21 source/setup.py
Baixar importâncias

@@ -10,3 +10,6 @@ for target, _ in targets_models.items():
 
 gdd.download_file_from_google_drive(file_id='1CLdiHuNq_QNOcv6fTs8HlOnlScE9RNCO',
                                     dest_path='sql/data.db', overwrite=True, showsize=True)
+
+gdd.download_file_from_google_drive(file_id='1lKhaK1dir8ywBHf22kbnGjTHuKsJ2-i4',
+                                    dest_path='model/importances.pkl', overwrite=True, showsize=True)

### 2021-06-21 paper/paper.tex
Gráficos

@@
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% How to use writeLaTeX:
+%
+% You edit the source code here on the left, and the preview on the
+% right shows you the result within a few seconds.
+%
+% Bookmark this page and share the URL with your co-authors. They can
+% edit at the same time!
+%
+% You can upload figures, bibliographies, custom classes and
+% styles using the files menu.
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \documentclass[12pt]{article}
@@
 \usepackage{graphicx,url}
+\usepackage{subcaption}
 
@@
     \email{paula.borba@ga.ita.br,  wallace.souza@ga.ita.br}
-%	\today
 }
@@
 
-Impacts of climate change on hydro resources have become a growing concern, since it may affect negatively the electricity supply in coming years, particularly in countries highly dependent on hydropower, like Brazil.
-Therefore, the ability to assess and make predictions about the water level and streamflow of reservoirs allows important decisions to be made by competent authorities in order to ensure that water resources can be appropriately allocated.
+\begin{figure}[htbp]
+    {\centering
+    \begin{subfigure}[b]{.49\textwidth}
+        \centering
+        \includesvg{graphs/R2_w01}
+        \caption{<caption>}
+    \end{subfigure}
+    \begin{subfigure}[b]{.49\textwidth}
+        \centering
+        \includesvg{graphs/R2_w15}
+        \caption{<caption>}
+    \end{subfigure}}
 
-Over 63\% of the whole electricity supply in Brazil came from hydropower in 2019.
-However, Brazil faces reduced availability of hydro resources in recent years.
-ONS affirmed that Brazil is experiencing a drought period similar to the period between 1948 to 1955 \cite{onsseca}, while other studies suggested the occurrence of more frequent and anomalous droughts in Brazil due to climate changes.
-
-% TODO: No final do paragrafo, talvez seja bom elencar os estudos que sugerem uma frequência anormal
-Climate change may affect hydroelectric production mainly because of the precipitation altering.
-Reduced precipitation may compromise river flows, while increased precipitation could render a power plant unproductive as it could exceed its capacity.
-Also, the increase in average temperature may alter the soil moisture level, interfering with the flow and storage of water in dams \cite{Mukheibir2013}.
-According to \cite{Lyra2018}, southeastern Brazil will suffer a significant reduction of precipitation during the twenty-first century.
-While \cite{Queiroz2016} noted that the Southeast may be benefited from climate changes with considerable increases in electricity generation, but in other regions of Brazil, climate change will likely cause the reduction of hydroelectricity production.
-
-Global Climate Models (GCMs) simulate interactions of climate systems based on physical principles and observational data.
-They are tools that are used to predict weather patterns and to explore climate sensitivities in scenarios of climate change \cite{gcm}.
-To provide long-term simulations of climate change at higher resolution, GCMs are downscaled to Regional Climate Models (RCMs), which include projections at a higher resolution.
-This study explored the impact of two climate change scenarios,  Representative Concentration Pathway (RCP 4.5 and RCP 8.5), on hydro resources in the Três Marias reservoir, located in the Southeast of Brazil.
-
-Thus, the objective of this article is to make a forecast of the flow and water level of a reservoir using machine learning predictive models.
-The dataset come from a simulation of the Platform Project of the Center for Weather Forecasting and Climate Studies (CPTEC) of the National Institute for Space Research (INPE).
-The simulated input data using will be (i) precipitation, (ii) evaporation, (iii) temperature, and (iv) surface runoff.
-As output data, the water level and stream flow of the reservoir.
-
-\section{Related bibliography}
-
-As pointed by \cite{ardabili2019deep}, using physical or statistical models to study hydrological processes can raise several problems, such as high computational costs, weakness in uncertainty analysis and the need for a large amount of comprehensive data.
-Therefore, the usage of machine learning methods have been growing among research community, as shown in Figure~\ref{fig:num_publications}, as a strategy to tackle some of these problems.
-In addition, several researchers have used machine learning to predict hydrological patterns, such as to predict flow in future hydrographic basins considering climate changes \cite{bhatta2019evaluation}, daily flow to river \cite{kassem2020predicting}, and flow of the hydroelectric plant reservoir \cite{essenfelder2020smart}.
-Table~\ref{tab:variables-literature} shows the variables used in these studies.
-
-\begin{figure}[htb]
+    \begin{subfigure}[b]{.49\textwidth}
         \centering
-    \begin{tikzpicture}
-        \begin{axis}[
-                width=13cm,
-                height=4.5cm,
-                style={/pgf/number format/set thousands separator={}},
-                axis x line=bottom,
-                axis y line=left,
-                axis line style={shorten <=-10},
-                xmin=2009,
-                xmax=2022,
-                xtick={2010,...,2021},
-                ymin=0,
-                ymax=120,
-                ytick={10, 30,...,110},
-                xlabel={Year},
-                xticklabel style={rotate=45},
-                ylabel={Number of publications}
-            ]
-            \addplot[black,ybar,fill, fill opacity=0.0, bar width = 0.8] table {
-                2021    34
-                2020    111
-                2019    75
-                2018    33
-                2017    16
-                2016    16
-                2015    5
-                2014    7
-                2013    6
-                2012    7
-                2011    3
-                2010    3
-            };
-        \end{axis}
-    \end{tikzpicture}
-    \caption{Number of publications about machine learning applied in the study of hydrological processes.
-The search string used was ``\texttt{(Machine learning) AND (hydrological))}'' in the Engineering Village database}
-    \label{fig:num_publications}
-\end{figure}
-
-\begin{table}[htbp]
-\centering
-\caption{Variables in the related bibliography}
-\label{tab:variables-literature}
-    \begin{tabulary}{\textwidth}{lCCC}
-        \toprule
-           & Kassem et al (2019) & Bhatta et al (2019) & Essenfelder et al (2020) \\
-        \midrule
-            Precipitation & X & X & X \\
-            Temperature & X & X & X \\
-            Relative humidity & X &  & \\
-            Wind speed  & X &  & \\
-            Landuse classes  &  & X & \\
-            Stream flow  & X & X & X \\
-        \bottomrule
-    \end{tabulary}
-\end{table}
-
-\cite{kicsi2007streamflow} presented a comparison between four different neural networks algorithms applied in short term daily stream flow forecast.
-They concluded that the Levenberg–Marquardt algorithm, when compared with backpro pagation, conjugate gradient and cascade correlation takes a relatively small time for training, while giving the best stream flow forecast, even using correlation analysis as the method to determine the input vector.
-Concurrently, \cite{petty2018streamflow} assessed the capacity of employing machine learning techniques and big data to impute streamflow using historical and real-time data.
-Their results demonstrated that applying Streamflow Hydrology Estimate using Machine Learning (SHEM) models could be used to estimate streamflow and even to help fill missing data from historical datasets.
-
-\section{Material and methods}
-Using two different scenarios, RCP 4.5 and RCP 8.5, we explored the long-term effect of climate variables on flow and water level variables.
-The predictor and outcome variables are described in Figure~\ref{fig:approach}.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.5\linewidth]{Figures/approach.png}
-  \caption{Methodology}
-  \label{fig:approach}
-\end{figure}
-
-\subsection{Descriptive analysis}
-We chose Três Marias reservoir for two main reasons.
-First because of its importance to the Brazilian electricity system, since this reservoir holds a hydropower plant of 396 MW of installed capacity.
-Second, there are available climate data at a high spatial resolution (5 km) for the Southeastern region, where Três Marias is located.
-Três Marias reservoir is inside the Alto São Francisco sub-basin in Minas Gerais state (Figure \ref{fig:studyarea}).
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.6\linewidth]{Figures/mapa_v0.png}
-  \caption{Alto São Franscisco sub-basin.}
-  \label{fig:studyarea}
-\end{figure}
-
-In order to reduce the complexity, we opted to use the contribution of each river and sub-basin area to the stream flow and water level from \cite{fundep}.
-Thus, we made an manual classification from all coordinates to 11 points based in the distance to the nearest river which represents the area of influence (Figure \ref{fig:coord}).
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.6\linewidth]{Figures/coord_contribuicao.png}
-  \caption{Coordinate points and their contribution rate.}
-  \label{fig:coord}
-\end{figure}
-
-\subsection{Data}
-In order to apply machine learning to forecast streamflow and water level for Três Marias reservoir, we used data generated by CPTEC/INPE and made available on the Projeta Platform \cite{chou2014assessment,chou2014evaluation,Lyra2018}.
-This data is split into historic observations, ranging from January, 1961 to December, 2005 and projections modeled according to RCP 4.5 and RCP 8.5, ranging from January, 2006 to December, 2099.
-For our study, since observations of flow and water levels for Três Marias reservoir made available by ONS \cite{onsnivel,onsvazao} start in 1999, historical data collected from the Projeta platform were filtered to start at the same period.
-
-A list of potential predictors is shown in Table~\ref{tab:predictors}.
-The study area, Alto São Francisco sub-basin, was subdivided in 1672 equally sized regions, and for each of these regions, data regarding temperature, precipitation, evaporation and surface runoff were gathered, according to temporal resolution, historical data and modeled projections shown in Table~\ref{tab:predictors}, which lists the potential predictors alongside to their respective measurement units.
-Additionally, Table~\ref{tab:outcomes} shows the measurement units, temporal resolution and date range for data regarding streamflow and water level for the Três Marias reservoir.
-
-\begin{table}[htbp]
-\centering
-\caption{Relation of predictors considered in this study obtained on the Projeta Platform \cite{chou2014assessment,chou2014evaluation,Lyra2018}}
-\label{tab:predictors}
-    \resizebox{0.8\columnwidth}{!}{%
-    \begin{tabulary}{\textwidth}{lcCcCr}
-        \toprule
-            Variable       & Unit & Temporal resolution & Historical data & Modeled Projections \\
-        \midrule
-            Temperature (T)    & ºC   & Daily               & 1999-2005       & 2006-2099 \\
-            Precipitation (P) & mm   & Daily               & 1999-2005       & 2006-2099 \\
-            Evaporation (E)    & mm   & Daily               & 1999-2005       & 2006-2099 \\
-            Surface runoff (R) & mm   & Daily               & 1999-2005       & 2006-2099 \\
-        \bottomrule
-    \end{tabulary}
-    }
-\end{table}
-
-\begin{table}[htbp]
-    \centering
-    \caption{Outcome variables made available by ONS \cite{onsnivel,onsvazao}}
-    \label{tab:outcomes}
-    \resizebox{0.6\columnwidth}{!}{%
-    \begin{tabular}{lccr}
-        \toprule
-            Variable       & Unit & Temporal resolution & Historical data \\
-        \midrule
-            Streamflow (SF)           & m³/s & Daily               & 1999-2021 \\
-            Water level (WL)    & m    & Daily               & 1999-2021 \\
-        \bottomrule
-    \end{tabular}
-    }
-    \label{tab:my_label}
-\end{table}
-
-%Descrever os dados: números de exemplos, de atributos, os tipos e escalas dos atributos, se é classificação/regressão, etc..
-Se for classificação, dizer a proporção de exemplos na classe com mais observações (majoritária).
-
-%\subsection{Técnicas}
-
-%\section{Results}\label{sec:figs}
-
-% Precisamos citar: Integração de dados (merges), Eliminação de atributos e como tratamos valores fora da faixa (RNOF) e limpeza de dados (verificar)
-
-\section{Data preprocessing}
-
-In order to organize the data into attribute-value format, we followed three main steps: data reduction, data integration and data transformation.
-Originally, we had a large dataset with the variables P, E, T and R related to 1673 coordinate points inside the sub-basin limits in the period between 1999 to 2005.
-
-We aggregated the coordinate points into 11 groups of rivers according to their contribution area to the Três Marias reservoir and removed information that are not necessary to the model, such as the latitude and longitude (Figure \ref{fig:step1}).
-
-Then, in order to reduce the original dataset and transform it into the information that represents all points in their respective group, we considered the daily sum for P and E, and the daily mean for T and R values (Figure \ref{fig:step2}).
-Thus, we created weather variables of P, E, T and R for 11 groups of rivers instead of 1673 points.
-
-Although time is an important feature relating to weather variability, we did not consider the date as an attribute, since the seasonal changes are already incorporated in the magnitude of the other variables.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=1\linewidth, trim=0cm 0 0 2.2cm,clip=true]{Figures/STEP1.png}
-  \caption{Step 1: data reduction.}
-  \label{fig:step1}
+        \includesvg{graphs/R2_w30}
+        \caption{<caption>}
+    \end{subfigure}
+    \caption{<caption>}
+    \label{<label>}
 \end{figure}
@@
 \begin{figure}[htbp]
+    {\centering
+    \begin{subfigure}[b]{.49\textwidth}
         \centering
-  \includegraphics[width=1\linewidth, trim=0cm 0 0 1.5cm,clip=true]{Figures/STEP2.png}
-  \caption{Step 2: feature aggregation.}
-  \label{fig:step2}
+        \includesvg{graphs/MAE_w01}
+        \caption{<caption>}
+    \end{subfigure}
+    \begin{subfigure}[b]{.49\textwidth}
+        \centering
+        \includesvg{graphs/MAE_w15}
+        \caption{<caption>}
+    \end{subfigure}}
+
+    \begin{subfigure}[b]{.49\textwidth}
+        \centering
+        \includesvg{graphs/MAE_w30}
+        \caption{<caption>}
+    \end{subfigure}
+    \caption{<caption>}
+    \label{<label1>}
 \end{figure}
-
-In data integration, we combined three different databases.
-Here, we joined weather variables (predictors) with Water Level (WL) and Stream-Flow (SF), which are both features related to the Três Marias reservoir (Figure \ref{fig:step3}).
-Besides that, they are the target variables.
-The mathematical representation of merging databases is illustrated in Figure \ref{fig:matriz1}.
-Finally, we obtained the data matrix in attribute-value format, which is represented in Figure \ref{fig:matriz2}.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=1\linewidth, trim=0cm 0 0 1.2cm,clip=true]{Figures/STEP3.png}
-  \caption{Step 3: data integration.}
-  \label{fig:step3}
-\end{figure}
-
-\begin{figure}[htbp]
-  \centering
-  \input{Data/matrix1}
-  %\includegraphics[width=1\linewidth]{Figures/matriz1.png}
-  \caption{Matrix of different databases}
-  \label{fig:matriz1}
-\end{figure}
-
-\begin{figure}[htbp]
-  \centering
-  \input{Data/matrix2}
-  %\includegraphics[width=1\linewidth]{Figures/matriz2.png}
-  \caption{Data matrix}
-  \label{fig:matriz2}
-\end{figure}
-
-\subsection{Normalization}
-The results of the earlier procedures indicate there's values in multiple scales in the dataset.
-Since the prediction model selected in the next steps can be sensitive to scale features, we normalized the data with the min-max normalization with the mathematical equation represented in \ref{eqn:Normalization}:
-\begin{equation}
-\label{eqn:Normalization}
-x' = \frac{x-min(x)}{max(x)-min(x)}
-\end{equation}
-where ~$x$ is the present value, ~$min(x)$ is the minimum value of the feature and ~$max(x)$ is the maximum value of the feature.
-
-We applied this procedure for every variable in the dataset.
-While only the scale in the values changed with the normalization procedure, it is easier to identify the variation in mean and standard deviation between variables, as shown in Figure \ref{fig:Normalized} for the inputs in the first river.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.6\linewidth, trim=0cm 0 0 .7cm,clip=true]{Figures/Normalização.png}
-  \caption{Boxplot of inputs from raw data x normalized data}
-  \label{fig:Normalized}
-\end{figure}
-
-Other preprocessing procedures, such as feature selection or dimensionality reduction, are not applicable since we filtered the data in the first step (download from Projeta Platform).
-Therefore, all data is composed of numeric values that make other transformation techniques unnecessary.
-
-In the future, we intend to verify if the variation between measures can generate better results for the prediction model than absolute values, such as comparison between reservoir levels and variation between days.
-
-In conclusion, we converted the database in the format attribute-value with normalized values between features.
-The "CSV" file extension archive in the annex provides further information about the newly generated dataset.
 

### 2021-06-21 model_load.py

@@ -1,75 +1,355 @@
 # %%
-import os.path
-
-import eli5
-import joblib
-import numpy
-import pandas
-from eli5.sklearn import PermutationImportance
-from google_drive_downloader import GoogleDriveDownloader as gdd
-from IPython import get_ipython
-from mlxtend.feature_selection import ColumnSelector
-from sklearn.compose import ColumnTransformer, TransformedTargetRegressor
-from sklearn.dummy import DummyRegressor
-from sklearn.ensemble import RandomForestRegressor, StackingRegressor
-from sklearn.impute import SimpleImputer
-from sklearn.inspection import permutation_importance
-from sklearn.linear_model import Ridge
-from sklearn.metrics import (accuracy_score, f1_score, make_scorer,
-                             precision_score, recall_score)
-from sklearn.model_selection import (GridSearchCV, KFold, cross_val_score,
-                                     cross_validate, train_test_split)
-from sklearn.pipeline import Pipeline
-from sklearn.preprocessing import MinMaxScaler, StandardScaler
-from sklearn.svm import SVR
-from sklearn.tree import DecisionTreeRegressor
-from sqlalchemy import select
-from sqlalchemy.sql.expression import true
-
-from source.custom_transfomers.debug_transformer import Debug
+from source.project_utils.data_manipulation import generate_aggregation
+from source.project_utils.constants import targets, targets_models
+from source.data_base.models import models
+from source.data_base.connection import session
 from source.custom_transfomers.time_window_transformer import \
     TimeWindowTransformer
-from source.data_base.connection import session
-from source.data_base.models import models
-from source.project_utils.constants import targets, targets_models
-from source.project_utils.data_manipulation import generate_aggregation
+from source.custom_transfomers.debug_transformer import Debug
+from sqlalchemy.sql.expression import label, true
+from sqlalchemy import select
+from sklearn.tree import DecisionTreeRegressor
+from sklearn.svm import SVR
+from sklearn.preprocessing import MinMaxScaler, StandardScaler
+from sklearn.pipeline import Pipeline
+from sklearn.model_selection import (GridSearchCV, KFold, cross_val_score,
+                                     cross_validate, train_test_split)
[…]

### 2021-06-21 source/project_utils/constants.py

@@ -11,5 +11,4 @@ targets_models = {
     'streamflow-s_d15': '1nLVWmqW8Uayrrq5xEX3flwH-bIPaaBJ9',
     'streamflow-s_d22': '1MikMxtvQAN_T8GEcRs-9sKBxsCoFYGAm',
     'streamflow-s_d29': '1wDVAzePrwtGlx9JY0zkowgNHv8cpodr2',
-    'streamflow-s_d30': '1gKCdINSDXNN9jDMidsjB7Iw4c2JtRNto',
 }

### 2021-06-22 source/custom_transfomers/reshape_transformer.py
Reshape trasnformer

@@ -1,0 +1,45 @@
+from typing import Any, Set, Type, Union, Tuple
+
+import numpy
+from pandas import DataFrame, MultiIndex, Index
+import pandas
+from sklearn.base import BaseEstimator, RegressorMixin, TransformerMixin
+from sklearn.utils.validation import check_array, check_is_fitted
+from sqlalchemy.sql.schema import RETAIN_SCHEMA
+from source.project_utils.data_manipulation import ColumnsLoc
+
+
+class ShapeTransformer(TransformerMixin, BaseEstimator):
+
+    def __init__(self, shape=(-1, 1)) -> None:
+        self.shape = shape
+
+    def fit(self, X: numpy.ndarray, y=None):
+
+        if hasattr(self, 'is_fitted_'):
+            del self.is_fitted_
+
+        if hasattr(self, 'is_ndarray_'):
+            del self.is_ndarray_
+            del self.shape_
+
+        self.is_fitted_ = True
+
+        if isinstance(X, numpy.ndarray):
+            self.is_ndarray_ = True
+            self.shape_ = X.shape
+
+        return self
+
+    def transform(self, X: numpy.ndarray, copy=None) -> numpy.ndarray:
+        check_is_fitted(self, 'is_fitted_')
+
+        if self.is_ndarray_:
+            if ShapeTransformer._is_oneD(self.shape_):
+                return X.reshape(self.shape)
+
+        return X
+
+    @staticmethod
+    def _is_oneD(shape: numpy.ndarray):
+        return len(shape) == 1

### 2021-06-23 model_load.py
Melhorias na construção do ScoresDataFrame

@@ -69,10 +69,14 @@ idx = pandas.IndexSlice
 
 # %%
 # Carregar o modelo
+debug = True
+models_path = 'model/'
+if debug:
+    models_path += 'debug/'
 regression_models = {}
 targets_models = range(1, 30, 7)
 for target in targets_models:
-    regression_models[target] = joblib.load(f'model/streamflow-s_d{target:02d}.pkl')
+    regression_models[target] = joblib.load(models_path + f'streamflow-s_d{target:02d}.pkl')
 
 # %%
 ImportancesDataFrame = joblib.load('model/importances.pkl')
@@ -88,11 +92,11 @@ def cm_to_inches(cm):
 columns_ = [
     (('shifted_level', ''), 'Shifted Level', 'b'),
     (('evaporation', 1), 'Evaporation River 1', 'g'),
-    (('evaporation', 6), 'evaporation River 6', 'r'),
-    (('evaporation', 7), 'evaporation River 7', 'c'),
-    (('evaporation', 10), 'evaporation River 10', 'k'),
-    (('month', ''), 'month', 'y'),
-    (('precipitation', 4), 'precipitation River 4', 'm'),
+    (('evaporation', 6), 'Evaporation River 6', 'r'),
+    (('evaporation', 7), 'Evaporation River 7', 'c'),
+    (('evaporation', 10), 'Evaporation River 10', 'k'),
+    (('month', ''), 'Month', 'y'),
+    (('precipitation', 4), 'Precipitation River 4', 'm'),
     (('shifted_streamflow', ''), 'Shifted Streamflow', 'brown'),
     (('surface_runoff', 10), 'Surface Runoff River 10', 'indianred'),
     (('surface_runoff', 11), 'Surface Runoff River 11', 'darkseagreen'),
@@ -117,19 +121,15 @@ def plot_save_importance(
                                          for row in ImportancesDataFrame.transpose().loc[idx[:, :, :], idx[:, windowing, 'mean']].values]
     aaaa = aaaa.sort_values(('mean', windowing, 'mean'), ascending=False)[:5]
     aaaa = aaaa.drop(columns=('mean', windowing, 'mean')).transpose()
-    for column in columns_:  # aaaa.loc[idx[:, windowing, 'mean']].columns:
+    for column in aaaa.loc[idx[:, windowing, 'mean']].columns:
 
-        if column[0] in aaaa.loc[idx[:, windowing, 'mean']].columns:
-            plt.plot(
[…]

### 2021-06-23 model_load.py
teste com novos parâmetros no gridsearch

@@ -69,7 +69,7 @@ idx = pandas.IndexSlice
 
 # %%
 # Carregar o modelo
-debug = True
+debug = False
 models_path = 'model/'
 if debug:
     models_path += 'debug/'
@@ -79,7 +79,7 @@ for target in targets_models:
     regression_models[target] = joblib.load(models_path + f'streamflow-s_d{target:02d}.pkl')
 
 # %%
-ImportancesDataFrame = joblib.load('model/importances.pkl')
+ImportancesDataFrame = joblib.load(models_path + 'importances.pkl')
 
 #
 # %%

### 2021-06-23 model_search.py

@@ -19,22 +19,17 @@ from sklearn.metrics import (accuracy_score, f1_score, make_scorer,
 from sklearn.model_selection import (GridSearchCV, KFold, cross_val_score,
                                      cross_validate, train_test_split)
 from sklearn.pipeline import Pipeline
-from sklearn.preprocessing import MinMaxScaler, OrdinalEncoder, StandardScaler
+from sklearn.preprocessing import MinMaxScaler, OrdinalEncoder, StandardScaler, OneHotEncoder
 from sklearn.svm import SVR
 from sklearn.tree import DecisionTreeRegressor
 from sqlalchemy import select
 from sqlalchemy.sql.expression import true
 
-from source.custom_transfomers.dataframe_transformers import (
-    AggregateTransformer,
-    GroupByTransformer,
-    PivotTransformer,
-    ResetIndexTransformer,
-    TestTargetColumnSelector,
-    TestTransformer)
+
 from source.custom_transfomers.debug_transformer import Debug, DebugY
 from source.custom_transfomers.time_window_transformer import \
     TimeWindowTransformer
+from source.custom_transfomers.reshape_transformer import ShapeTransformer
 from source.data_base.connection import session
 from source.data_base.models import models
 from source.project_utils.constants import targets
@@ -129,32 +124,45 @@ date_cols = ['month']
 shifted_cols = ['shifted_level', 'shifted_streamflow']
 
 # %%
-cv = KFold(n_splits=10, random_state=seed, shuffle=True)
-
-
 pipeline_steps = [
     ('columns', ColumnTransformer([
         (
-            'normalize',
-            MinMaxScaler(feature_range=(0, 1)),
-            ColumnsLoc(var_cols).get_loc
-        ),
-        (
             'month',
-            OrdinalEncoder(),
+            'passthrough',
             ColumnsLoc(date_cols).get_loc
         ),
         (
-            'historical',
+            'vars',
             'passthrough',
-            ColumnsLoc(shifted_cols).get_loc
-        )
+            ColumnsLoc(var_cols + shifted_cols).get_loc
+        ),
     ],
         remainder='drop',
     )),
     ('reg', DummyRegressor())
 ]
 
+# %%
+svr_month_pipeline_steps = [
+    ('reshape', ShapeTransformer()),
+    ('scale', 'passthrough'),
+]
+
+# %%
+svr_pipeline_steps = [
+    ('columns', ColumnTransformer([
+        (
+            'month',
[…]

### 2021-06-23 model_search.py
Salvando RawDataFrame num arquivo

@@ -1,4 +1,7 @@
 # %%
+import multiprocessing
+import os
+
 import eli5
 import joblib
 import numpy
@@ -19,17 +22,17 @@ from sklearn.metrics import (accuracy_score, f1_score, make_scorer,
 from sklearn.model_selection import (GridSearchCV, KFold, cross_val_score,
                                      cross_validate, train_test_split)
 from sklearn.pipeline import Pipeline
-from sklearn.preprocessing import MinMaxScaler, OrdinalEncoder, StandardScaler, OneHotEncoder
+from sklearn.preprocessing import (MinMaxScaler, OneHotEncoder, OrdinalEncoder,
+                                   StandardScaler)
 from sklearn.svm import SVR
 from sklearn.tree import DecisionTreeRegressor
 from sqlalchemy import select
 from sqlalchemy.sql.expression import true
 
-
 from source.custom_transfomers.debug_transformer import Debug, DebugY
+from source.custom_transfomers.reshape_transformer import ShapeTransformer
 from source.custom_transfomers.time_window_transformer import \
     TimeWindowTransformer
-from source.custom_transfomers.reshape_transformer import ShapeTransformer
 from source.data_base.connection import session
 from source.data_base.models import models
 from source.project_utils.constants import targets
@@ -49,21 +52,28 @@ pandas.set_option('display.max_columns', 51)
 # %%
 # Construção do dataframe utilizando buscas no banco de dados sql
 print('SQL query...')
-query = select(
-    models.Variables.date,
-    models.Variables.precipitation.label('precipitation'),
-    models.Variables.temperature.label('temperature'),
-    models.Variables.evaporation.label('evaporation'),
-    models.Variables.surface_runoff.label('surface_runoff'),
-    models.Coordinate.river_id.label('river'),
-    models.Reservoir.level,
-    models.Reservoir.streamflow
-).\
-    join(models.Variables.coordinate).\
-    join(models.Reservoir, models.Variables.date == models.Reservoir.date).\
-    order_by(models.Variables.date)
 
-RawDataFrame = pandas.read_sql(query, session.bind)
+if not os.path.isfile('raw_data_frame.pkl'):
+    query = select(
+        models.Variables.date,
+        models.Variables.precipitation.label('precipitation'),
+        models.Variables.temperature.label('temperature'),
[…]

### 2021-06-23 model_load.py
Função para criar a label das colunas no plot de importância

@@ -1,4 +1,5 @@
 # %%
+from typing import Tuple, Union
 from source.project_utils.data_manipulation import generate_aggregation
 from source.project_utils.constants import targets, targets_models
 from source.data_base.models import models
@@ -42,37 +43,14 @@ plt.style.use('default')
 
 idx = pandas.IndexSlice
 
-# matplotlib.use("pgf")
-# matplotlib.rcParams.update({
-#     "pgf.texsystem": "pdflatex",
-#     'font.family': 'serif',
-#     'text.usetex': True,
-#     'pgf.rcfonts': False,
-# })
-
-# # %%
-# query = select(
-#     models.Variables.date,
-#     models.Variables.precipitation.label('precipitation'),
-#     models.Variables.temperature.label('temperature'),
-#     models.Variables.evaporation.label('evaporation'),
-#     models.Variables.surface_runoff.label('surface_runoff'),
-#     models.Coordinate.river_id.label('river'),
-#     models.Reservoir.level,
-#     models.Reservoir.streamflow
-# ).\
-#     join(models.Variables.coordinate).\
-#     join(models.Reservoir, models.Variables.date == models.Reservoir.date).\
-#     order_by(models.Variables.date)
-
-# RawDataFrame = pandas.read_sql(query, session.bind)
-
 # %%
 # Carregar o modelo
-debug = False
+debug = True
 models_path = 'model/'
 if debug:
     models_path += 'debug/'
+
+# %%
 regression_models = {}
 targets_models = range(1, 30, 7)
 for target in targets_models:
@@ -89,20 +67,12 @@ def cm_to_inches(cm):
     return cm / 2.54
 
 
-columns_ = [
-    (('shifted_level', ''), 'Shifted Level', 'b'),
-    (('evaporation', 1), 'Evaporation River 1', 'g'),
-    (('evaporation', 6), 'Evaporation River 6', 'r'),
-    (('evaporation', 7), 'Evaporation River 7', 'c'),
-    (('evaporation', 10), 'Evaporation River 10', 'k'),
-    (('month', ''), 'Month', 'y'),
-    (('precipitation', 4), 'Precipitation River 4', 'm'),
-    (('shifted_streamflow', ''), 'Shifted Streamflow', 'brown'),
-    (('surface_runoff', 10), 'Surface Runoff River 10', 'indianred'),
-    (('surface_runoff', 11), 'Surface Runoff River 11', 'darkseagreen'),
-]
[…]

### 2021-06-23 model_search.py
Renomeado de roll para windowing

@@ -253,7 +253,7 @@ grid_search_params = dict(
         'neg_mean_absolute_percentage_error'
     ],
     cv=KFold(n_splits=10, shuffle=True, random_state=seed),
-    n_jobs=multiprocessing.cpu_count() - 1,
+    n_jobs=-1,
     verbose=1,
     error_score="raise",
     refit='r2'
@@ -273,7 +273,7 @@ agg.update(runoff_agg)
 # %%
 print('GridSearch...')
 shift_regressors = {}
-for shift in range(1, 30, 7):
+for shift in range(15, 30, 7):
     ShiftedDataFrame = ConsolidatedDataFrame.copy()
 
     ShiftedDataFrame.insert(
@@ -299,6 +299,7 @@ for shift in range(1, 30, 7):
             windowed_data={}
         )
         for roll in [1, 15, 30]:
+            print(f"{shift} - {roll} - {target}")
             search = GridSearchCV(Pipeline(steps=pipeline_steps), **grid_search_params)
             WindowedDataFrame = TimeWindowTransformer(var_cols, roll, agg, True).fit_transform(ShiftedDataFrame)
             target_regressor[target]['estimators'][roll] = search.fit(WindowedDataFrame, WindowedDataFrame[target])

### 2021-06-24 model_search.py
Remoção do cálculo das importâncias

@@ -273,7 +273,7 @@ agg.update(runoff_agg)
 # %%
 print('GridSearch...')
 shift_regressors = {}
-for shift in range(15, 30, 7):
+for shift in range(1, 30, 7):
     ShiftedDataFrame = ConsolidatedDataFrame.copy()
 
     ShiftedDataFrame.insert(
@@ -298,46 +298,19 @@ for shift in range(15, 30, 7):
             estimators={},
             windowed_data={}
         )
-        for roll in [1, 15, 30]:
-            print(f"{shift} - {roll} - {target}")
+        for windowing in [1, 15, 30]:
+            print(f"{shift} - {windowing} - {target}", end=' - ')
             search = GridSearchCV(Pipeline(steps=pipeline_steps), **grid_search_params)
-            WindowedDataFrame = TimeWindowTransformer(var_cols, roll, agg, True).fit_transform(ShiftedDataFrame)
-            target_regressor[target]['estimators'][roll] = search.fit(WindowedDataFrame, WindowedDataFrame[target])
-            target_regressor[target]['windowed_data'][roll] = WindowedDataFrame
+            WindowedDataFrame = TimeWindowTransformer(var_cols, windowing, agg, True).fit_transform(ShiftedDataFrame)
+            target_regressor[target]['estimators'][windowing] = search.fit(WindowedDataFrame, WindowedDataFrame[target])
+            target_regressor[target]['windowed_data'][windowing] = WindowedDataFrame
             if search.best_score_ > target_regressor[target]['best_score']:
                 target_regressor[target]['best_score'] = search.best_score_
                 target_regressor[target]['best_params'] = search.best_params_
                 target_regressor[target]['best_estimator'] = search.best_estimator_
-                target_regressor[target]['best_windowing'] = roll
+                target_regressor[target]['best_windowing'] = windowing
                 target_regressor[target]['best_windowed_data'] = WindowedDataFrame
 
         joblib.dump(target_regressor[target], f'model/debug/{target}-s_d{shift:02d}.pkl')
-
-# %%
-# Testar importância das features
-ImportancesDataFrame = pandas.DataFrame()
-names = ['target', 'windowing', 'fun']
-columns = shift_regressors[1]['streamflow']['best_windowed_data'].columns
-for lagging in [1, 8, 15, 22, 29]:
-    regression_models = shift_regressors[lagging]
-    for target in ['streamflow']:
-        for windowing in [1, 15, 30]:
-            importances = permutation_importance(
[…]

### 2021-06-24 model_load.py
Pequenas alterações nos gráficos e no ScoresDataFrame

@@ -204,7 +204,7 @@ for windowing, lagging in multiIndex:
         UniqueRegressorsDataFrame['name'] = names
 
         ScoresDataFrame.loc[(windowing, lagging), (regressor, score, statistic)] = \
-            UniqueRegressorsDataFrame[UniqueRegressorsDataFrame['name'] == regressor][scores[score][statistic]].values
+            UniqueRegressorsDataFrame[UniqueRegressorsDataFrame['name'] == regressor][scores[score][statistic]].values[0]
 
 
 # %%
@@ -225,9 +225,11 @@ def plot_save_score(
         save_folder += '/'
 
     _, ax = plt.subplots()
-    n_plots = len(ScoresDataFrame.loc[idx[windowing], idx[:, scorer]].columns.unique('regressor'))
+
+    regressors = ScoresDataFrame.loc[idx[windowing], idx[:, scorer]].columns.unique('regressor')
+    n_plots = len(regressors)
     k = -0.2 * n_plots / 2
-    for reg in ScoresDataFrame.loc[idx[windowing], idx[:, scorer]].columns.unique('regressor'):
+    for reg in regressors:
         trans1 = Affine2D().translate(k, 0.0) + ax.transData
         plt.errorbar(
             ScoresDataFrame.loc[windowing][(reg, scorer, 'mean')].index,
@@ -254,7 +256,7 @@ width = 7.3
 # %%
 plot_save_score(windowing=1,
                 scorer='R2',
-                # lim=(0.2, 1),
+                lim=(0.5, 1),
                 legend_loc=None,
                 x_label='Shift (days)',
                 y_label='$R^2$ score',
@@ -265,7 +267,7 @@ plot_save_score(windowing=1,
 # %%
 plot_save_score(windowing=15,
                 scorer='R2',
-                # lim=(0.2, 1),
+                lim=(0.5, 1),
                 legend_loc=None,
                 x_label='Shift (days)',
                 y_label='$R^2$ score',
@@ -276,7 +278,7 @@ plot_save_score(windowing=15,
 # %%
 plot_save_score(windowing=30,
                 scorer='R2',
-                # lim=(0.2, 1),
+                lim=(0.5, 1),
                 legend_loc='upper left',
                 x_label='Shift (days)',
                 y_label='$R^2$ score',
@@ -287,6 +289,7 @@ plot_save_score(windowing=30,
 # %%
 plot_save_score(windowing=1,
                 scorer='MAE',
+                lim=(-300, -80),
                 legend_loc=None,
[…]

### 2021-06-24 model_load.py
Debug = false

@@ -45,7 +45,7 @@ idx = pandas.IndexSlice
 
 # %%
 # Carregar o modelo
-debug = True
+debug = False
 models_path = 'model/'
 if debug:
     models_path += 'debug/'

### 2021-06-24 paper/paper.tex
Paper com novos gráficos

@@
 \usepackage{tabulary}
+
+\usepackage{caption}
+\usepackage{subcaption}
+\captionsetup{subrefformat=parens}
 \usepackage{svg}
@@
 \usepackage{subcaption}
-
 %\usepackage[brazil]{babel}
@@
 
-\title{Assessment of climate change impacts on hydro resources in Três Marias reservoir using machine learning models}
+\title{Machine learning methods in streamflow prediction  considering climate variable windowing}
 
@@
 
-%\begin{resumo}
+\begin{abstract}
+    Hydropower operation and planning requires streamflow forecasts.
+Machine learning methods combined with atmospheric climate variables complement physical models, particularly, where in-situ observations are not available.
+To make predictions of streamflow of a reservoir, we test four ML techniques: Support Vector Regressor (SVR), Simple Decision Tree, Random Forest (RF), and Stacking Regressor, which combines SVR, RF and Ridge.
+Also, we add prediction horizon varying from 0 to 29 days for streamflow, and windowing of 1, 15, and 30 days for atmospheric climate variables.
+Results indicate a better fit in the RF model.
+Models outperform when a shorter horizon prediction of streamflow and longer windowing for climate variables are considered.
 
-%\end{resumo}
+\end{abstract}
 
@@
 
+Over 63\% of the whole electricity supply in Brazil came from hydropower in 2019.
+However, Brazil faces reduced availability of hydro resources in recent years \cite{onsseca}.
+Factors as changes in rainfall-runoff and temperature patterns affect the physical process of streamflow and water level in reservoirs.
+Impacts of climate variability on hydro resources have become a growing concern, since it may affect negatively the electricity supply in coming years, particularly in countries highly dependent on hydropower, like Brazil.
+Therefore, the ability to assess and make predictions about the streamflow of reservoirs allow important decisions to be made by competent authorities in order to ensure that water resources can be appropriately allocated.
+
+Machine learning methods have been developed as a complement to physical hydrologic models, particularly, in contexts where data are limited to support those physical-based models.
+Further, the development of satellite data associated with in situ observations to monitor climate and hydrology makes available data of climate and streamflow more widely in data-poor regions \cite{julie}.
+
+In this work, we make short-term forecast streamflow using Três Marias reservoir as study case.
+We use atmospheric variables as precipitation, evaporation, temperature, and surface runoff, available in Projeta project \cite{inpe}, and observed data of streamflow and water level of Três Marias reservoir \cite{onsnivel,onsvazao}.
+We compared four machine learning methods in order to access the best performance of short-term predictions.
+Moreover, we evaluate the impact of adding a lag of a certain period in streamflow and water level data, as well as, windowing the climate variables.
+
+In Section 2, the related bibliography regarding machine learning models in the context of streamflow is introduced.
+The study area and data used for the modeling are presented in Section 3, while the pre-processing data and results are described in Section 4, followed by a analysis about the results in Section 5 and conclusion in Section 6.
+
+\section{Related bibliography}
+
+Using physical or statistical models to study hydrological processes can raise several problems, such as high computational costs, weakness in uncertainty analysis and the need for a large number of comprehensive data \cite{ardabili2019deep}.
+Therefore, as a strategy to tackle some of these problems, machine learning methods, particularly Support Vector for Regression (SVR) and Random Forest (RF), have been growing in the water resource research community.
+
+Support Vector for Regression (SVR) has an appropriate performance in solving small-scale, over-learning and nonlinear high-dimensional problems in the field of hydrological prediction \cite{xu,wang}.
+This regression model can be constructed using a nonlinear mapping function onto a high-dimensional space, in which the non-linear separable problem becomes linearly separable in space \cite{maity}.
+However, this model is highly dependent on kernel parameters.
+
+Conventional decision tree learners generate a tree by splitting the data based on the values of predictors.
+Each node in the tree represents a test of some attribute, while each branch from that node is the possible value for that attribute.
+
+Random Forest (RF) is also a widely used algorithm because of its satisfactory predictive power for time series \cite{wang}.
+Additionally, the model has the advantage of a simplified data pre-processing \cite{li}.
+The RF model combines a simple decision tree, in which each tree is generated by selecting random samples, and features from all predictors \cite{naghibi}.
+
+According to \cite{dou}, ensemble machine learning techniques
+provide an accurate and efficient solution in spatial modeling.
+Stacking ensemble ML algorithm learns how to combine the predictors from other well-performing models.
+In our case, we combined both SVR, RF models and Ridge regressors.
+
+The natural complexity of hydrologic variables drives the approaches of mathematical models to actual output, in this case, streamflow.
+Although the selection of input variables varies according to the study, precipitation and temperature are common variables as observed in \cite{kassem2020predicting,essenfelder2020smart,bhatta2019evaluation}.
+Historical data of streamflow itself as input variable was present in \cite{li}.
+In order to approach the natural structure of a reservoir, we included in our work the variables runoff and evaporation, besides precipitation, evaporation, streamflow and water level.
+
+\section{Material and methods}
+
+We selected Três Marias reservoir as a case study to forecast its short-term streamflow using four machine learning techniques: SVR, decision tree, RF and Stacking Regressor, combining both SVR and RF.
+We first describe the study area and data collection, then, we presented the parameters and applied methods of machine learning.
+
+\subsection{Study area and descriptive analysis}
+We chose Três Marias reservoir for two main reasons.
+First because of its importance to the Brazilian electricity system, since this reservoir holds a hydropower plant of 396 MW of installed capacity.
+Second, there are available atmospheric climate data at a high spatial resolution (5 km) for the Southeastern region, where Três Marias is located.
+Três Marias reservoir is inside the Alto São Francisco sub-basin in Minas Gerais state (Figure \ref{fig:studyarea}).
+
 \begin{figure}[htbp]
-    {\centering
+    \centering
+    \includegraphics[width=0.9\linewidth]{Figures/mapa.pdf}
+    \caption{Alto São Francisco sub-basin.}
+    \label{fig:studyarea}
+\end{figure}
+
+We defined the river-group of each coordinate point using as reference \cite{fundep}.
+
+\subsection{Data}
+
+We collected atmospheric climate data generated by CPTEC/INPE, available on the Projeta Platform \cite{chou2014assessment,chou2014evaluation,Lyra2018}, to forecast streamflow for Três Marias reservoir.
+The available data includes the period from January 1999 to December 2005.
+We combined that with observed data of streamflow and water level for the same period, available on \cite{onsnivel,onsvazao}.
+In our study, streamflow is the target variable, but its previous value is also input to the prediction.
+All variables are shown in Table~\ref{tab:predictors}.
+For further information of data exploration, see Appendix A.
+
+\begin{table}[htbp]
+    \centering
+    \caption{Relation of predictors.}
+    \label{tab:predictors}
+    %\resizebox{0.6\columnwidth}{!}{%
+    \renewcommand{\arraystretch}{.8}
+    \small\begin{tabulary}{\textwidth}{lcCcCr}
+        \toprule
+        \normalsize\bfseries{Variable} &
+        \normalsize\bfseries{Unit} &
+        \normalsize\bfseries{Temporal resolution} &
+        \normalsize\bfseries{Historical data}  \\
+        \midrule
+        Temperature (T)    & ºC   & Daily   & 1999-2005 \\
+        Precipitation (P)  & mm   & Daily   & 1999-2005 \\
+        Evaporation (E)    & mm   & Daily   & 1999-2005 \\
+        Surface runoff (R) & mm   & Daily   & 1999-2005 \\
+        Streamflow (SF)    & m³/s & Daily   & 1999-2005 \\
+        Water level (WL)   & m    & Daily   & 1999-2005 \\
+        Month (M)          & -   & -        & 1999-2005 \\
+        \bottomrule
+    \end{tabulary}
+    %}
+\end{table}
+
+%Descrever os dados: números de exemplos, de atributos, os tipos e escalas dos atributos, se é classificação/regressão, etc..
+Se for classificação, dizer a proporção de exemplos na classe com mais observações (majoritária).
+
+\subsection{Methods and evaluation criteria}
+
+We tested four models: SVR, simple decision tree, RF and stacking ensemble method.
+We evaluated the performance of each model adding a lag for streamflow and water level varying from 1, 8, 15, 22 and 29 days.
+Furthermore, we included a windowing for climate variables in intervals of 1, 15, and 30 days.
+
+Using the python library Scikit-Learn \cite{scikit-learn}, we applied iterative steps, also known as pipelines, to detail the parameters and regression techniques.
+After these steps, we analyse the predictors results by the metrics R-squared score and Mean Absolute Error (MAE) to analyse their performance.
+Figure \ref{fig:pipelines} describes the pipelines.
+
+We used a combination of parameters to tune the model with R-squared score as metric, leveraging the GridSearchCV scikit-learn tool.
+In GridSearchCV, we define some constraints of each parameter, available in Appendix B.
+
+\begin{figure}[htbp]
+    \centering
+    {\footnotesize\includesvg[width=.8\linewidth]{Diagrams/gridsearch.drawio}}
+    \caption{Pipelines in the machine learning models.}
+    \label{fig:pipelines}
+\end{figure}
+
+To evaluate the performance of SVR, RF, Stacking Regressor and decision tree, we use two criteria: Mean Absolute Error (MAE) and R-squared.
+MAE represents the average of the absolute difference between the actual and predicted values.
+This procedure is represented by the Equation
+\begin{equation}
+    \label{eqn:mae}
+    MAE={\frac{1}{n}\sum_{i-1}^{n}|(X_{pred}-X_{i})|}
+\end{equation}
+in which $n$ means the number of non-missing data, $i$ is the variable, $X_{pred}$ and $X_{i}$ mean the predicted and the actual value, respectively.
+
+Conversely, $R^2$ represents the proportion of the variance in the dependent variable which is explained by the regression model, given by the Equation
+\begin{equation}
+    \label{eqn:r2}
+    R^2={1 - \frac{\sum_{i=1}^{n}(X_{pred}-X_{i})^{2}}{\sum_{i=1}^{n}(X_{i}-X_{mean})^{2}}}
+\end{equation}
+where $X_{mean}$ is $\frac{1}{n}\sum_{i=1}^{n}X_{i}$ and $X_{pred}$ and $X_{i}$ are the predicted and it's corresponding true value, while $n$ is the size of the test sample.
+R-square is a indicator of goodness, indicating how well the model can predict unseen samples.
+The best possible R-squared value is $1.0$ and it can be negative, as the model gets arbitrarily worse.
+I
+
+\section{Data preprocessing}
+
+We followed three main steps to organize the data into attribute-value format: data reduction, data integration, and data transformation,  Originally, we had a large dataset with the variables P, E, T, and R related to 1673 coordinate points inside the sub-basin limits in the period between 1999 to 2005.
+
+We aggregated the coordinate points into 11 groups of rivers according to their contribution area to the Três Marias reservoir.
+We considered the daily sum for P and E, and the daily mean for T and R values.
+Thus, we created weather variables of P, E, T, and R for 11 groups of rivers instead of 1673 points (see Appendix).
+Additinally, we considered month as a qualitative attribute in order to capture seasonal changes in streamflow.
+
+\subsection{Normalization}
+
+The variables were normalized to prevent the model from being dominated by large values. We used the Min-max normalization given by the Equation
+\begin{equation}
+    \label{eqn:Normalization}
+    x' = \frac{x-min(x)}{max(x)-min(x)}
+\end{equation}
+to adjust the scale of the inputs.
+Where the value $x$ is the present value, $min(x)$ is the minimum value of the feature, and $max(x)$ is the maximum value of the feature.
+Then, we transformed that to their original scale.
+
+We applied this procedure for every variable in the dataset, except for month.
+While only the scale in the values changed with the normalization procedure, it is easier to identify the variation in mean and standard deviation between variables, as shown in Figure \ref{fig:Normalized} for the inputs in the first river.
+
+\begin{figure}[htbp]
+    \centering
+    \includegraphics[width=0.8\linewidth, trim=0cm 0 0 .7cm,clip=true]{Figures/Normalização.png}
+    \caption{Boxplot of inputs from raw data x normalized data}
+    \label{fig:Normalized}
+\end{figure}
+
+Other preprocessing procedures, such as feature selection or dimensionality reduction, are not applicable since we filtered the data in the first step (download from Projeta Platform).
+Therefore, all data is composed of numeric values that make other transformation techniques unnecessary.
+
+\section{Results and discussion}
+
+Having tested horizon prediction of streamflow and water level varying from 1 to 29 days, and windowing at intervals of 1, 15 and 30 days for the climate variables, results indicated better fit for horizon of one day and window of 30 days.
+However, the models presented different results when they are compared to each other, as shown in Figure \ref{fig:r2}.
+
+\begin{figure}[htbp]
+    \centering
     \begin{subfigure}[b]{.49\textwidth}
         \centering
-        \includesvg{graphs/R2_w01}
-        \caption{<caption>}
+        \includesvg{Graphs/R2_w01}
+        \captionsetup{margin={.15\textwidth, 0cm}}
+        \caption{1 day windowing}
+        \label{fig:r2_1}
     \end{subfigure}
@@
         \centering
-        \includesvg{graphs/R2_w15}
-        \caption{<caption>}
-    \end{subfigure}}
+        \includesvg{Graphs/R2_w15}
+        \captionsetup{margin={.15\textwidth, 0cm}}
+        \caption{15 days windowing}
+        \label{fig:r2_15}
+    \end{subfigure}
+    \begin{subfigure}[b]{\textwidth}
+        \centering
+        \includesvg{Graphs/R2_w30}
+        \captionsetup{justification=justified,singlelinecheck=false}
+        \captionsetup{margin={.23\linewidth, 0cm}}
+        \caption{30 days windowing}
+        \label{fig:r2_30}
+    \end{subfigure}
+    \caption{\textbf{$R^2$} for \subref{fig:r2_1} 1, \subref{fig:r2_15} 15, and \subref{fig:r2_30} 30 days windowing}
+    \label{fig:r2}
+\end{figure}
 
+According to the increase of streamflow  and water level horizon, SVR decreased its R-squared in all considered periods of windowing.
+However, other models presented an unexpected performance varying the R-squared score in different ways.
+Regarding the 1 day-windowing, RF, Stacking Regressor and decision tree models had a turning point in 15 days-lagging, in which the performance reached its lowest point and rose again.
+This trend is kept in 15 days-windowing for Stacking Regressor, but for RF and decision tree, the performance with the horizon increase tends to establish.
+In 30 days-windowing, the models presented a greater $R^2$ variation in the interval between 0 to 7 days of streamflow and water level prediction horizon
+and then, that variation became less accentuated, except for SVR.
+
+Results indicated that the variable shifted streamflow lost its strength with the horizon increasing, while the models become more dependent on other variables, as indicated in Figure \ref{fig:sfig1_relevance}, \ref{fig:sfig2_relevance}, and \ref{fig:sfig3_relevance}.
+Month variable had its importance increased in 1 day and 15 days of window, while in 30 days-windowing it is decreased.
+Evaporation starts being more relevant with 30-days windowing.
+
+\begin{figure}[htbp]
+    \centering
+    \includesvg{Graphs/Importance_w01}
+    \caption{Variable relevance: 1 day-windowing.}
+    \label{fig:sfig1_relevance}
+\end{figure}
+\begin{figure}[htbp]
+    \centering
+    \includesvg{Graphs/Importance_w15}
+    \caption{Variable relevance: 15 days-windowing.}
+    \label{fig:sfig2_relevance}
+\end{figure}
+\begin{figure}[htbp]
+    \centering
+    \includesvg{Graphs/Importance_w30}
+    \caption{Variable relevance: 30 days-windowing.}
+    \label{fig:sfig3_relevance}
+\end{figure}
+
+Figure \ref{fig:sfigMAE} presents MAE values for different period-windowing.
+MAE were increased with  higher horizon, except for RF model in 15 and 30 days-windowing, which had lower MAE with a greater prediction horizon.
+However, if the error are compared analyzing the period-windowing, results indicate that there is a slight improvement regards MAE when this period is increased.
+This suggests a better performance when climate data are analyzed within a greater windowing.
+In a physical process, that might represent the period of delay that climate variables, which were collected in the large sub-basin extension, have an impact on the upstream reservoir.
+
+\begin{figure}[htbp]
+    \centering
     \begin{subfigure}[b]{.49\textwidth}
         \centering
-        \includesvg{graphs/R2_w30}
-        \caption{<caption>}
-    \end{subfigure}
-    \caption{<caption>}
-    \label{<label>}
-\end{figure}
-
-\begin{figure}[htbp]
-    {\centering
-    \begin{subfigure}[b]{.49\textwidth}
-        \centering
-        \includesvg{graphs/MAE_w01}
-        \caption{<caption>}
+        \includesvg{Graphs/MAE_w01}
+        \captionsetup{margin={.15\textwidth, 0cm}}
+        \caption{1 day windowing}
+        \label{fig:sfig1MAE}
     \end{subfigure}
@@
         \centering
-        \includesvg{graphs/MAE_w15}
-        \caption{<caption>}
-    \end{subfigure}}
+        \includesvg{Graphs/MAE_w15}
+        \captionsetup{margin={.15\textwidth, 0cm}}
+        \caption{15 days windowing}
+        \label{fig:sfig2MAE}
+    \end{subfigure}
+    \begin{subfigure}[b]{\textwidth}
+        \centering
+        \includesvg{Graphs/MAE_w30}
+        \captionsetup{margin={.23\linewidth, 0cm}}
+        \captionsetup{justification=justified,singlelinecheck=false}
+        \caption{30 days windowing}
+        \label{fig:sfig3MAE}
+    \end{subfigure}
+    \caption{Mean Absolute Errors (MAE) for \subref{fig:sfig1MAE} 1, \subref{fig:sfig2MAE} 15, and \subref{fig:sfig3MAE} 30 days windowing}
+    \label{fig:sfigMAE}
+\end{figure}
 
-    \begin{subfigure}[b]{.49\textwidth}
-        \centering
-        \includesvg{graphs/MAE_w30}
-        \caption{<caption>}
-    \end{subfigure}
-    \caption{<caption>}
-    \label{<label1>}
-\end{figure}
+RF had a better results in terms of MAE and $R^2$ if compared to other models.
+SVR presented presented worst results in both metrics, while both Stacking Regressor and Decision Tree had similar results.
+
+\section{Conclusion}
+
+In this work, we tested four methods to forecast streamflow using Três Marias reservoir as a study case.
+Results indicated that RF had the best fit over the models.
+For one-day prediction horizon for streamflow and water level , models presented R-squared values in the range 0.7 to 0.9, but that decreases according to the increase of  prediction horizon.
+With regards to windowing of climate variables, the best fits occur when considering a 30 days-windowing, which might be related to the delay of the physical process impact of all climate variables in the entire sub-basin on the reservoir upstream.
+
+SVR had the highest values of MAE, while RF had the lowest values.
+The Stacking Regressor and decision tree produced similar values of MAE.
+Surprisingly, RF presented a slight increase of MAE for greater days of lagging, and for the case of 15 and 30 days-windowing, while the other models presented a trend of having a decline.
+
+Empirical results obtained from these models showed the performance not only over the models, but also changing the lag of streamflow and water level, and the window period of climate variables.
+That is important in solving water resources management problems and improve the accuracy of flow prediction.
+
+Finally, other properties that have a direct impact on streamflow, as land-use, soil classification, and wind speed, might improve the approach of modeling to the actual physical phenomena.
+Thus, we suggest the inclusion of other climate variables, and also other reservoir study cases.
 
@@
 
+\appendix
+\input{Appendix}
+
 \end{document}

### 2021-06-24 model_load.py
Limpeza dos imports

@@ -1,41 +1,10 @@
 # %%
 from typing import Tuple, Union
-from source.project_utils.data_manipulation import generate_aggregation
-from source.project_utils.constants import targets, targets_models
-from source.data_base.models import models
-from source.data_base.connection import session
-from source.custom_transfomers.time_window_transformer import \
-    TimeWindowTransformer
-from source.custom_transfomers.debug_transformer import Debug
-from sqlalchemy.sql.expression import label, true
-from sqlalchemy import select
-from sklearn.tree import DecisionTreeRegressor
-from sklearn.svm import SVR
-from sklearn.preprocessing import MinMaxScaler, StandardScaler
-from sklearn.pipeline import Pipeline
-from sklearn.model_selection import (GridSearchCV, KFold, cross_val_score,
-                                     cross_validate, train_test_split)
-from sklearn.metrics import (accuracy_score, f1_score, make_scorer,
-                             precision_score, recall_score)
-from sklearn.linear_model import Ridge
-from sklearn.inspection import permutation_importance
-from sklearn.impute import SimpleImputer
-from sklearn.ensemble import RandomForestRegressor, StackingRegressor
-from sklearn.dummy import DummyRegressor
-from sklearn.compose import ColumnTransformer, TransformedTargetRegressor
-from mlxtend.feature_selection import ColumnSelector
-from IPython import get_ipython
-from google_drive_downloader import GoogleDriveDownloader as gdd
-from eli5.sklearn import PermutationImportance
-import tikzplotlib
+from source.project_utils.constants import targets_models
 import pandas
-import numpy
-import matplotlib
 import matplotlib.pyplot as plt
 import joblib
-import eli5
 from matplotlib.transforms import Affine2D
-from matplotlib import transforms
 
 pandas.set_option('display.max_columns', 51)
 

### 2021-06-24 model_search.py

@@ -1,52 +1,29 @@
 # %%
-import multiprocessing
 import os
 
-import eli5
 import joblib
-import numpy
 import pandas
-from eli5.sklearn import PermutationImportance
-from IPython import get_ipython
-from mlxtend.feature_selection import ColumnSelector
-from scipy.stats.stats import DescribeResult
-from sklearn.compose import (ColumnTransformer, TransformedTargetRegressor,
-                             make_column_selector)
+from sklearn.compose import ColumnTransformer, TransformedTargetRegressor
 from sklearn.dummy import DummyRegressor
 from sklearn.ensemble import RandomForestRegressor, StackingRegressor
-from sklearn.impute import SimpleImputer
-from sklearn.inspection import permutation_importance
 from sklearn.linear_model import Ridge
-from sklearn.metrics import (accuracy_score, f1_score, make_scorer,
-                             precision_score, recall_score)
-from sklearn.model_selection import (GridSearchCV, KFold, cross_val_score,
-                                     cross_validate, train_test_split)
+from sklearn.metrics import accuracy_score, make_scorer
+from sklearn.model_selection import GridSearchCV, KFold
 from sklearn.pipeline import Pipeline
-from sklearn.preprocessing import (MinMaxScaler, OneHotEncoder, OrdinalEncoder,
-                                   StandardScaler)
+from sklearn.preprocessing import MinMaxScaler, OneHotEncoder
 from sklearn.svm import SVR
 from sklearn.tree import DecisionTreeRegressor
 from sqlalchemy import select
-from sqlalchemy.sql.expression import true
 
-from source.custom_transfomers.debug_transformer import Debug, DebugY
 from source.custom_transfomers.reshape_transformer import ShapeTransformer
 from source.custom_transfomers.time_window_transformer import \
     TimeWindowTransformer
 from source.data_base.connection import session
 from source.data_base.models import models
-from source.project_utils.constants import targets
 from source.project_utils.data_manipulation import (ColumnsLoc,
                                                     generate_aggregation)
 
 pandas.set_option('display.max_columns', 51)
-
-# To add a new cell, type '# %%'
-# To add a new markdown cell, type '# %% [markdown]'
-
-# # %%
-# get_ipython().run_line_magic('load_ext', 'autoreload')
-# get_ipython().run_line_magic('autoreload', '2')
 
 
 # %%

### 2021-06-24 model_load.py
Geração de gráficos adicionais"

@@ -87,6 +87,7 @@ height = 5
 
 # %%
 plot_save_importance(windowing=1,
+                     lim=(-.2, 2.2),
                      legend_loc='upper left',
                      x_label='Shift (days)',
                      y_label='Importance',
@@ -96,6 +97,7 @@ plot_save_importance(windowing=1,
 
 # %%
 plot_save_importance(windowing=15,
+                     lim=(-.2, 2.2),
                      legend_loc='upper left',
                      x_label='Shift (days)',
                      y_label='Importance',
@@ -105,6 +107,7 @@ plot_save_importance(windowing=15,
 
 # %%
 plot_save_importance(windowing=30,
+                     lim=(-.2, 2.2),
                      legend_loc='upper left',
                      x_label='Shift (days)',
                      y_label='Importance',
@@ -126,12 +129,22 @@ scores = {
     'MAE': {
         'mean': 'mean_test_neg_mean_absolute_error',
         'std': 'std_test_neg_mean_absolute_error',
-        'rank': 'rank_test_neg_mean_absolute_error'
+        'rank': 'rank_test_neg_mean_absolute_error',
     },
     'R2': {
         'mean': 'mean_test_r2',
         'std': 'std_test_r2',
-        'rank': 'rank_test_r2'
+        'rank': 'rank_test_r2',
+    },
+    'MAPE': {
+        'mean': 'mean_test_neg_mean_absolute_percentage_error',
+        'std': 'std_test_neg_mean_absolute_percentage_error',
+        'rank': 'rank_test_neg_mean_absolute_percentage_error',
+    },
+    'RMSE': {
+        'mean': 'mean_test_neg_root_mean_squared_error',
+        'std': 'std_test_neg_root_mean_squared_error',
+        'rank': 'rank_test_neg_root_mean_squared_error',
     },
 }
 
@@ -284,6 +297,66 @@ plot_save_score(windowing=30,
                 legend_loc='upper left',
                 x_label='Shift (days)',
                 y_label='MAE score',
+                width=width + 6.5,
+                height=height,
+                save_folder='paper/graphs')
+
+# %%
+plot_save_score(windowing=1,
+                scorer='MAPE',
+                legend_loc=None,
+                x_label='Shift (days)',
+                y_label='MAPE score',
+                width=width,
+                height=height,
+                save_folder='paper/graphs')
+
+# %%
+plot_save_score(windowing=15,
+                scorer='MAPE',
+                legend_loc=None,
[…]

### 2021-06-24 paper/paper.tex
Extensão do parágrafo sobre o pipeline

@@
 We compared four machine learning methods in order to access the best performance of short-term predictions.
-Moreover, we evaluate the impact of adding a lag of a certain period in streamflow and water level data, as well as, windowing the climate variables.
+Moreover, we evaluate the impact of adding a horizon prediction of streamflow, as well as, windowing the climate variables.
 
-In Section 2, the related bibliography regarding machine learning models in the context of streamflow is introduced.
+In Section 2, the related bibliography regarding machine learning models in the context of streamflow prediction is introduced.
 The study area and data used for the modeling are presented in Section 3, while the pre-processing data and results are described in Section 4, followed by a analysis about the results in Section 5 and conclusion in Section 6.
@@
 
-We selected Três Marias reservoir as a case study to forecast its short-term streamflow using four machine learning techniques: SVR, decision tree, RF and Stacking Regressor, combining both SVR and RF.
+We selected Três Marias reservoir as a case study to forecast its short-term streamflow using four machine learning techniques: SVR, decision tree, RF and Stacking Regressor, combining SVR, RF and Ridge.
 We first describe the study area and data collection, then, we presented the parameters and applied methods of machine learning.
@@
 Três Marias reservoir is inside the Alto São Francisco sub-basin in Minas Gerais state (Figure \ref{fig:studyarea}).
+We defined the river-group of each coordinate point using as reference \cite{fundep}.
 
@@
 \end{figure}
-
-We defined the river-group of each coordinate point using as reference \cite{fundep}.
 
@@
 We tested four models: SVR, simple decision tree, RF and stacking ensemble method.
-We evaluated the performance of each model adding a lag for streamflow and water level varying from 1, 8, 15, 22 and 29 days.
-Furthermore, we included a windowing for climate variables in intervals of 1, 15, and 30 days.
+We evaluated the performance of each model adding a windowing for climate variables in intervals of 1, 15, and 30 days.
+Furthermore, we tested horizon predictions varying from 0 to 29 days.
 
  Using the python library Scikit-Learn \cite{scikit-learn}, we applied iterative steps, also known as pipelines, to detail the parameters and regression techniques.
-After these steps, we analyse the predictors results by the metrics R-squared score and Mean Absolute Error (MAE) to analyse their performance.
-Figure \ref{fig:pipelines} describes the pipelines.
+After these steps, we analyse the predictors results by the metrics R-squared score and Mean Absolute Error (MAE) to assess their performance.
+As shown in Figure \ref{fig:pipelines}, the developed pipeline starts by applying a time window transformation on precipitation, evaporation, temperature and surface runoff variables.
+This transformation consisted in aggregating data from a specified time window, i.e., adding precipitation from the last fifteen or thirty days.
+Additionally, when the regressor was a SVR, data were normalized, as described in Section \ref{sec:DataTransformation}.
 
@@
   \centering
+%   \includegraphics[width=1\linewidth]{Diagrams/Pipelines.PNG}
   {\footnotesize\includesvg[width=.8\linewidth]{Diagrams/gridsearch.drawio}}
@@
 MAE represents the average of the absolute difference between the actual and predicted values.
+We included MAE as an error metric because of it provides an interpretation of error on the same scale of streamflow.
+
 This procedure is represented by the Equation
@@
 
-We followed three main steps to organize the data into attribute-value format: data reduction, data integration, and data transformation,  Originally, we had a large dataset with the variables P, E, T, and R related to 1673 coordinate points inside the sub-basin limits in the period between 1999 to 2005.
+We followed three main steps to organize the data into attribute-value format: data reduction, data integration, and data transformation.
+Other preprocessing procedures, such as feature selection, are not applicable since we filtered the data directly from Projeta Platform.
+
+\subsection{Data reduction and integration}
+
+Originally, we had a large dataset with the variables P, E, T, and R related to 1673 coordinate points inside the sub-basin limits in the period between 1999 to 2005.
 
@@
 We considered the daily sum for P and E, and the daily mean for T and R values.
-Thus, we created weather variables of P, E, T, and R for 11 groups of rivers instead of 1673 points (see Appendix).
-Additinally, we considered month as a qualitative attribute in order to capture seasonal changes in streamflow.
+Thus, we created weather variables of P, E, T, and R for 11 groups of rivers instead of 1673 points.
+Additionally, we considered month as a qualitative attribute in order to capture seasonal changes in streamflow.
 
-\subsection{Normalization}
+\subsection{Data transformation} \label{sec:DataTransformation}
 
-The variables were normalized to prevent the model from being dominated by large values. We used the Min-max normalization given by the Equation
+All variables, except month variable, were normalized when evaluating regressors that are sensitive to data scale, in order to prevent them from being dominated by variables with higher order of magnitude.
+We used the Min-Max normalization given by the Equation
     \begin{equation}
@@
 
-We applied this procedure for every variable in the dataset, except for month.
 While only the scale in the values changed with the normalization procedure, it is easier to identify the variation in mean and standard deviation between variables, as shown in Figure \ref{fig:Normalized} for the inputs in the first river.
@@
   \includegraphics[width=0.8\linewidth, trim=0cm 0 0 .7cm,clip=true]{Figures/Normalização.png}
-    \caption{Boxplot of inputs from raw data x normalized data}
+  \caption{Boxplot of inputs from normalized data (left) and raw data (rifht)}
   \label{fig:Normalized}
@@
 
-Other preprocessing procedures, such as feature selection or dimensionality reduction, are not applicable since we filtered the data in the first step (download from Projeta Platform).
-Therefore, all data is composed of numeric values that make other transformation techniques unnecessary.
+For month variable we apply one-hot encoding, which is often used for categorical variables.
 
@@
 Having tested horizon prediction of streamflow and water level varying from 1 to 29 days, and windowing at intervals of 1, 15 and 30 days for the climate variables, results indicated better fit for horizon of one day and window of 30 days.
-However, the models presented different results when they are compared to each other, as shown in Figure \ref{fig:r2}.
+However, the models presented different results of R\textsuperscript{2} when they are compared to each other, as shown in Figures \ref{fig:r2_1}, \ref{fig:r2_15} and \ref{fig:r2_30}.
 
@@
     \end{subfigure}
-    \caption{\textbf{$R^2$} for \subref{fig:r2_1} 1, \subref{fig:r2_15} 15, and \subref{fig:r2_30} 30 days windowing}
+    \caption{R\textsuperscript{2} for \subref{fig:r2_1} 1, \subref{fig:r2_15} 15, and \subref{fig:r2_30} 30 days windowing}
     \label{fig:r2}
@@
 
-According to the increase of streamflow  and water level horizon, SVR decreased its R-squared in all considered periods of windowing.
-However, other models presented an unexpected performance varying the R-squared score in different ways.
-Regarding the 1 day-windowing, RF, Stacking Regressor and decision tree models had a turning point in 15 days-lagging, in which the performance reached its lowest point and rose again.
-This trend is kept in 15 days-windowing for Stacking Regressor, but for RF and decision tree, the performance with the horizon increase tends to establish.
-In 30 days-windowing, the models presented a greater $R^2$ variation in the interval between 0 to 7 days of streamflow and water level prediction horizon
-and then, that variation became less accentuated, except for SVR.
+According to the increase of prediction horizon of streamflow up to 8 days, the R\textsuperscript{2} had an accentuated decrease in all models in 1 day-windowing.
+This trend is kept in 15 and 30 days windowing only for Decision Tree, while for SVR and Stacking Regressor the decrease is smaller and for Random Forest, R\textsuperscript{2} presented a growth.
+
+After 8 days of prediction horizon, the R\textsuperscript{2} tends to be constant in 15 and 30 days-windowing cases.
+In 1 day-windowing case, the models presented an unexpected performance, since the models had a turning point in 15 days-horizon, in which the performance reached its lowest point and rose again.
+
+Random forest outperforms in all considered cases of prediction horizon and windowing in terms of R\textsuperscript{2}.
+Decision Tree Regressor has the worst fit, while SVR and Stacking Regressor had similar results, although Stacking Regressor slightly presents higher values of R\textsuperscript{2}.
 
 Results indicated that the variable shifted streamflow lost its strength with the horizon increasing, while the models become more dependent on other variables, as indicated in Figure \ref{fig:sfig1_relevance}, \ref{fig:sfig2_relevance}, and \ref{fig:sfig3_relevance}.
-Month variable had its importance increased in 1 day and 15 days of window, while in 30 days-windowing it is decreased.
-Evaporation starts being more relevant with 30-days windowing.
+The importance of month variable increases in 1 day-windowing until prediction horizon of 22 days, then starts decreasing.
+In 15 days-windowing, month variable importance increases until 15 days, has a reduction in 22 days, and then grows again.
+In the case of 30 days-windowing, evaporation becomes the most relevant variable.
 
@@
   \end{figure}
+
 \begin{figure}[htbp]
@@
 
-Figure \ref{fig:sfigMAE} presents MAE values for different period-windowing.
-MAE were increased with  higher horizon, except for RF model in 15 and 30 days-windowing, which had lower MAE with a greater prediction horizon.
-However, if the error are compared analyzing the period-windowing, results indicate that there is a slight improvement regards MAE when this period is increased.
+Figure \ref{fig:sfigMAE} presents MAE values for different period-windowing and indicate how big of an error we can expect from the forecast on average.
+In all cases and models, the MAE values were greater than 100 m\textsuperscript{3}/s.
+However, MAE is higher for all models in 1 day-windowing case.
+In 15 and 30-days windowing, MAE was similar to each other, and greatly reduced if compared to 1 day-windowing.
 This suggests a better performance when climate data are analyzed within a greater windowing.
 In a physical process, that might represent the period of delay that climate variables, which were collected in the large sub-basin extension, have an impact on the upstream reservoir.
+
+Analyzing the prediction horizon, results indicate that MAE tends to be constant for SVR and Decision Tree after 8 days of the horizon.
+MAE in RF model decreases slightly with the horizon time prediction expansion in 15 and 30 days-windowing.
 
@@
 
-RF had a better results in terms of MAE and $R^2$ if compared to other models.
-SVR presented presented worst results in both metrics, while both Stacking Regressor and Decision Tree had similar results.
+In general terms, RF had a better results of MAE, if compared to other models. SVR presented the worst result, while both Stacking Regressor and Decision Tree had similar results.
 
@@
 Empirical results obtained from these models showed the performance not only over the models, but also changing the lag of streamflow and water level, and the window period of climate variables.
-That is important in solving water resources management problems and improve the accuracy of flow prediction.
+That is important in solving water resources management problems and improve the streamflow prediction.
 

### 2021-06-25 model_search.py
100 nas random forest

@@ -182,7 +182,7 @@ grid_search_params = dict(
                 None
             ],
             'reg__regressor__random_state': [seed],
-            'reg__regressor__n_estimators': [10, 250],
+            'reg__regressor__n_estimators': [10, 100, 250],
             'columns__vars': [
                 MinMaxScaler(feature_range=(0, 1)),
                 'passthrough'
@@ -203,7 +203,7 @@ grid_search_params = dict(
                     final_estimator=Ridge(random_state=seed)
                 )
             ],
-            'reg__RandomForest__regressor__n_estimators': [10, 250],
+            'reg__RandomForest__regressor__n_estimators': [10, 100, 250],
             'reg__RandomForest__transformer': [
                 MinMaxScaler(feature_range=(0, 1)),
                 None

### 2021-06-25 model_load.py
Gráficos de predito x real

@@ -1,10 +1,18 @@
 # %%
+import datetime
 from typing import Tuple, Union
-from source.project_utils.constants import targets_models
-import pandas
-import matplotlib.pyplot as plt
+
 import joblib
+import matplotlib.pyplot as plt
+import numpy as np
+import pandas
+import scipy.interpolate
+import scipy.stats
 from matplotlib.transforms import Affine2D
+from sklearn.model_selection import cross_val_predict
+from sqlalchemy.sql.expression import label
+
+from source.project_utils.constants import targets_models
 
 pandas.set_option('display.max_columns', 51)
 
@@ -186,7 +194,8 @@ for windowing, lagging in multiIndex:
         UniqueRegressorsDataFrame['name'] = names
 
         ScoresDataFrame.loc[(windowing, lagging), (regressor, score, statistic)] = \
-            UniqueRegressorsDataFrame[UniqueRegressorsDataFrame['name'] == regressor][scores[score][statistic]].values[0]
+            UniqueRegressorsDataFrame[UniqueRegressorsDataFrame['name']
+                                      == regressor][scores[score][statistic]].values[0]
 
 
 # %%
@@ -238,7 +247,7 @@ width = 7.3
 # %%
 plot_save_score(windowing=1,
                 scorer='R2',
-                lim=(0.5, 1),
+                lim=(0.4, 1),
                 legend_loc=None,
                 x_label='Shift (days)',
                 y_label='$R^2$ score',
@@ -249,7 +258,7 @@ plot_save_score(windowing=1,
 # %%
 plot_save_score(windowing=15,
                 scorer='R2',
-                lim=(0.5, 1),
+                lim=(0.4, 1),
                 legend_loc=None,
                 x_label='Shift (days)',
                 y_label='$R^2$ score',
@@ -260,7 +269,7 @@ plot_save_score(windowing=15,
 # %%
 plot_save_score(windowing=30,
                 scorer='R2',
-                lim=(0.5, 1),
+                lim=(0.4, 1),
                 legend_loc='upper left',
                 x_label='Shift (days)',
                 y_label='$R^2$ score',
@@ -274,7 +283,7 @@ plot_save_score(windowing=1,
                 lim=(-300, -80),
                 legend_loc=None,
                 x_label='Shift (days)',
-                y_label='MAE score',
+                y_label='Negative MAE score',
                 width=width,
                 height=height,
                 save_folder='paper/graphs')
[…]

### 2021-06-25 paper/paper.tex
Sobre k-fold

@@
 \begin{abstract}
-Hydropower operation and planning requires streamflow forecasts.
+Hydropower operation and planning require streamflow forecasts.
 Machine learning methods combined with atmospheric climate variables complement physical models, particularly, where in-situ observations are not available.
-To make predictions of streamflow of a reservoir, we test four ML techniques: Support Vector Regressor (SVR), Simple Decision Tree, Random Forest (RF), and Stacking Regressor, which combines SVR, RF and Ridge.
+To make predictions of streamflow of a reservoir, we test four ML techniques: Support Vector Regressor (SVR), Simple Decision Tree, Random Forest (RF), and Stacking Regressor, which combines SVR, RF, and Ridge.
 Also, we add prediction horizon varying from 0 to 29 days for streamflow, and windowing of 1, 15, and 30 days for atmospheric climate variables.
@@
 As shown in Figure \ref{fig:pipelines}, the developed pipeline starts by applying a time window transformation on precipitation, evaporation, temperature and surface runoff variables.
-This transformation consisted in aggregating data from a specified time window, i.e., adding precipitation from the last fifteen or thirty days.
-Additionally, when the regressor was a SVR, data were normalized, as described in Section \ref{sec:DataTransformation}.
+This transformation consists in aggregating data from a specified time window, i.e., adding precipitation from the last fifteen or thirty days.
+Additionally, when evaluating SVR, data were normalized, as described in Section \ref{sec:DataTransformation}.
 
-We used a combination of parameters to tune the model with R-squared score as metric, leveraging the GridSearchCV scikit-learn tool.
+We used a combination of parameters to tune the model with an R-squared score as a metric, leveraging the GridSearchCV scikit-learn tool.
 In GridSearchCV, we define some constraints of each parameter, available in Appendix B.
+We adopted a cross-validation strategy to optimize these parameters and minimize the risk of overfitting.
+This was made to avoid testing regressor parameters on the same data it was trained for.
+The adopted approach was k-fold cross-validation, where the data set was divided into k groups - $k$ was set to 10 in this study - the regressors were trained in $k-1$ of these groups, and the fold left was used to test.
+As a consequence, the test was made using data the regressor have never seen.
+We calculated the final score to be the mean of the scores given by testing with each group.
 
@@
 MAE represents the average of the absolute difference between the actual and predicted values.
-We included MAE as an error metric because of it provides an interpretation of error on the same scale of streamflow.
+We included MAE as an error metric because it provides an interpretation of error on the same scale of streamflow.
 
@@
         \label{eqn:mae}
-        MAE={\frac{1}{n}\sum_{i-1}^{n}|(X_{pred}-X_{i})|}
+        MAE={\frac{1}{n}\sum_{i=0}^{n-1}|(X_{pred}-X_{i})|}
     \end{equation}
@@
         \label{eqn:r2}
-        R^2={1 - \frac{\sum_{i=1}^{n}(X_{pred}-X_{i})^{2}}{\sum_{i=1}^{n}(X_{i}-X_{mean})^{2}}}
+        R^2={1 - \frac{\sum\limits_{i=0}^{n-1}(X_{pred}-X_{i})^{2}}
+                      {\sum\limits_{i=0}^{n-1}(X_{i}-X_{mean})^{2}}}
     \end{equation}
-where $X_{mean}$ is $\frac{1}{n}\sum_{i=1}^{n}X_{i}$ and $X_{pred}$ and $X_{i}$ are the predicted and it's corresponding true value, while $n$ is the size of the test sample.
+where $X_{mean}$ is $\frac{1}{n}\sum_{i=0}^{n-1}X_{i}$ and $X_{pred}$ and $X_{i}$ are the predicted and it's corresponding true value, while $n$ is the size of the test sample.
 R-square is a indicator of goodness, indicating how well the model can predict unseen samples.
 The best possible R-squared value is $1.0$ and it can be negative, as the model gets arbitrarily worse.
-I
 

### 2021-06-25 model_load.py
Definição do layout dos gráficos de comparação

@@ -371,9 +371,9 @@ plot_save_score(windowing=30,
                 save_folder='paper/graphs')
 
 # %%
-date_start = datetime.date(2001, 8, 25)
-date_end = datetime.date(2001, 10, 30)
-interpolate_steps = 1
+date_start = datetime.date(2000, 1, 1)
+date_end = datetime.date(2005, 1, 1)
+interpolate_steps = 5
 time_slice = slice(date_start, date_end)
 time_slice_interpolate = slice(date_start, date_end, interpolate_steps)
 
@@ -383,10 +383,12 @@ predict_X_y_spline = scipy.interpolate.make_interp_spline(
     original_data.loc[time_slice_interpolate].index.values,
     original_data.loc[time_slice_interpolate].values
 )
-interpolated_original_y = predict_X_y_spline(original_data.loc[time_slice].index)
+interpolated_original_y = predict_X_y_spline(original_data.loc[time_slice].index).clip(min=0)
 
 # %%
-for shift in range(1, 30, 7):
+width = 15
+height = 5
+for shift in [1, 29]:  # range(1, 30, 7):
     cv_predict = cross_val_predict(
         regression_models[shift]['best_estimator'],
         regression_models[shift]['best_windowed_data'],
@@ -403,9 +405,11 @@ for shift in range(1, 30, 7):
         predicted_values.loc[time_slice_interpolate].values
     )
 
-    interpolated_predicted_y = predict_X_y_spline(predicted_values.loc[time_slice].index)
-    plt.plot(original_data.loc[time_slice].index.values, interpolated_original_y,)
-    plt.plot(predicted_values.loc[time_slice].index, interpolated_predicted_y, label=shift)
+    interpolated_predicted_y = predict_X_y_spline(predicted_values.loc[time_slice].index).clip(min=0)
+    plt.gcf().set_size_inches(cm_to_inches(width), cm_to_inches(height))
+    plt.plot(original_data.loc[time_slice].index.values, interpolated_original_y, label='True')
+    plt.plot(predicted_values.loc[time_slice].index, interpolated_predicted_y, label='Predicted')
 
-    plt.legend()
+    plt.legend(loc='upper right')
+    plt.savefig(f'paper/Graphs/Comparisson_d{shift:02d}.svg')
     plt.show()

### 2021-06-25 model_load.py
Gráficos de importâncias categorizadas

@@ -45,6 +45,9 @@ def cm_to_inches(cm):
 
 
 def column_label(column: Tuple[str, Union[str, int]]):
+    if isinstance(column, str) == 1:
+        return column.replace('_', ' ').capitalize()
+
     label = column[0].replace('_', ' ').capitalize()
     if column[1] != "":
         label += f' River {column[1]:02d}'
@@ -91,7 +94,7 @@ def plot_save_importance(
 
 # %%
 width = 15
-height = 5
+height = 4.5
 
 # %%
 plot_save_importance(windowing=1,
@@ -122,6 +125,75 @@ plot_save_importance(windowing=30,
                      width=width,
                      height=height,
                      save_folder='paper/graphs')
+
+
+# %%
+def plot_save_categorized_importance(
+        windowing=1,
+        lim=None,
+        legend_loc=None,
+        save_folder: str = '.',
+        x_label=None,
+        y_label=None,
+        width=6.4,
+        height=4.8):
+    if not save_folder.endswith('/'):
+        save_folder += '/'
+    aaaa = ImportancesDataFrame.copy().transpose()
+    aaaa = aaaa.groupby(level=0).sum()[(aaaa.groupby(level=0).sum().T != 0).any()]
+    aaaa = aaaa.sort_values((1, windowing, 'mean'), ascending=False).T
+    for column in aaaa.loc[idx[:, windowing, 'mean']].columns.sort_values():
+
+        plt.plot(
+            aaaa.loc[idx[:, windowing, 'mean']][column].index,
+            aaaa.loc[idx[:, windowing, 'mean']][column],
+            label=column_label(column),
+        )
+
+    plt.ylim(lim)
+    plt.ylabel(y_label)
+    plt.xlabel(x_label)
+    plt.xticks(ImportancesDataFrame.loc[idx[:, windowing, 'mean']][column].index)
+
+    if legend_loc is not None:
+        plt.legend(loc=legend_loc, bbox_to_anchor=(1.05, 1))
+    plt.gcf().set_size_inches(cm_to_inches(width), cm_to_inches(height))
+    plt.grid(True)
+    plt.tight_layout()
+    plt.savefig(save_folder + f'CategorizedImportance_w{windowing:02d}.svg')
+
+
+# %%
+width = 7.3
+height = 5
+
+# %%
+plot_save_categorized_importance(windowing=1,
+                                 lim=(-0.2, 2.6),
+                                 x_label='Shift (days)',
+                                 y_label='Importance',
+                                 width=width,
[…]

### 2021-06-26 paper/paper.tex
Definição do layout dos gráficos de comparação
Gráficos de importâncias substituidos por gráficos de importâncias categorizadas

@@
 
+\begin{figure}[htbp]
+    \centering
+  %   \includegraphics[width=1\linewidth]{Diagrams/Pipelines.PNG}
+    {\small\includesvg[width=.8\linewidth]{Diagrams/gridsearch.drawio}}
+    \caption{Pipeline setup for machine learning model search}
+    \label{fig:pipelines}
+  \end{figure}
+
 We used a combination of parameters to tune the model with an R-squared score as a metric, leveraging the GridSearchCV scikit-learn tool.
@@
 The adopted approach was k-fold cross-validation, where the data set was divided into k groups - $k$ was set to 10 in this study - the regressors were trained in $k-1$ of these groups, and the fold left was used to test.
-As a consequence, the test was made using data the regressor have never seen.
-We calculated the final score to be the mean of the scores given by testing with each group.
-
-\begin{figure}[htbp]
-  \centering
-%   \includegraphics[width=1\linewidth]{Diagrams/Pipelines.PNG}
-  {\footnotesize\includesvg[width=.8\linewidth]{Diagrams/gridsearch.drawio}}
-  \caption{Pipelines in the machine learning models.}
-  \label{fig:pipelines}
-\end{figure}
+As a consequence, the test was made using data the regressor have never seen while training.
+We calculate the final score as the average of the scores obtained for each test group.
 
@@
 We included MAE as an error metric because it provides an interpretation of error on the same scale of streamflow.
-
 This procedure is represented by the Equation
@@
 in which $n$ means the number of non-missing data, $i$ is the variable, $X_{pred}$ and $X_{i}$ mean the predicted and the actual value, respectively.
-
 Conversely, $R^2$ represents the proportion of the variance in the dependent variable which is explained by the regression model, given by the Equation
@@
 
-After 8 days of prediction horizon, the R\textsuperscript{2} tends to be constant in 15 and 30 days-windowing cases.
-In 1 day-windowing case, the models presented an unexpected performance, since the models had a turning point in 15 days-horizon, in which the performance reached its lowest point and rose again.
-
-Random forest outperforms in all considered cases of prediction horizon and windowing in terms of R\textsuperscript{2}.
-Decision Tree Regressor has the worst fit, while SVR and Stacking Regressor had similar results, although Stacking Regressor slightly presents higher values of R\textsuperscript{2}.
-
-Results indicated that the variable shifted streamflow lost its strength with the horizon increasing, while the models become more dependent on other variables, as indicated in Figure \ref{fig:sfig1_relevance}, \ref{fig:sfig2_relevance}, and \ref{fig:sfig3_relevance}.
-The importance of month variable increases in 1 day-windowing until prediction horizon of 22 days, then starts decreasing.
-In 15 days-windowing, month variable importance increases until 15 days, has a reduction in 22 days, and then grows again.
-In the case of 30 days-windowing, evaporation becomes the most relevant variable.
-
-\begin{figure}[htbp]
-  \centering
-  \includesvg{Graphs/Importance_w01}
-  \caption{Variable relevance: 1 day-windowing.}
-  \label{fig:sfig1_relevance}
-  \end{figure}
-
-\begin{figure}[htbp]
-  \centering
-  \includesvg{Graphs/Importance_w15}
-  \caption{Variable relevance: 15 days-windowing.}
-  \label{fig:sfig2_relevance}
-  \end{figure}
-
-\begin{figure}[htbp]
-  \centering
-  \includesvg{Graphs/Importance_w30}
-  \caption{Variable relevance: 30 days-windowing.}
-  \label{fig:sfig3_relevance}
-\end{figure}
-
 Figure \ref{fig:sfigMAE} presents MAE values for different period-windowing and indicate how big of an error we can expect from the forecast on average.
@@
 In a physical process, that might represent the period of delay that climate variables, which were collected in the large sub-basin extension, have an impact on the upstream reservoir.
-
-Analyzing the prediction horizon, results indicate that MAE tends to be constant for SVR and Decision Tree after 8 days of the horizon.
-MAE in RF model decreases slightly with the horizon time prediction expansion in 15 and 30 days-windowing.
 
@@
 
+Analyzing the prediction horizon, results indicate that MAE tends to be constant for SVR and Decision Tree after 8 days of the horizon.
+MAE in RF model decreases slightly with the horizon time prediction expansion in 15 and 30 days-windowing.
+
+After 8 days of prediction horizon, the R\textsuperscript{2} tends to be constant in 15 and 30 days-windowing cases.
+In 1 day-windowing case, the models presented an unexpected performance, since the models had a turning point in 15 days-horizon, in which the performance reached its lowest point and rose again.
+
+Random forest outperforms in all considered cases of prediction horizon and windowing in terms of R\textsuperscript{2}.
+Decision Tree Regressor has the worst fit, while SVR and Stacking Regressor had similar results, although Stacking Regressor slightly presents higher values of R\textsuperscript{2}.
+
+Results indicated that the variable shifted streamflow lost its strength with the horizon increasing, while the models become more dependent on other variables, as indicated in Figure \ref{fig:CategorizedImportance}.
+The importance of month variable increases in 1 day-windowing until prediction horizon of 22 days, then starts decreasing.
+In 15 days-windowing, month variable importance increases until 15 days, has a reduction in 22 days, and then grows again.
+In the case of 30 days-windowing, evaporation becomes the most relevant variable.
+
+\begin{figure}[htbp]
+    \centering
+    \begin{subfigure}[b]{.49\textwidth}
+        \centering
+        \includesvg{Graphs/CategorizedImportance_w01}
+        \captionsetup{margin={.15\textwidth, 0cm}}
+        \caption{1 day windowing}
+        \label{fig:CategorizedImportance01}
+    \end{subfigure}
+    \begin{subfigure}[b]{.49\textwidth}
+        \centering
+        \includesvg{Graphs/CategorizedImportance_w15}
+        \captionsetup{margin={.15\textwidth, 0cm}}
+        \caption{15 days windowing}
+        \label{fig:CategorizedImportance02}
+    \end{subfigure}
+    \begin{subfigure}[b]{\textwidth}
+        \centering
+        \includesvg{Graphs/CategorizedImportance_w30}
+        \captionsetup{margin={.23\linewidth, 0cm}}
+        \captionsetup{justification=justified,singlelinecheck=false}
+        \caption{30 days windowing}
+        \label{fig:CategorizedImportance03}
+    \end{subfigure}
+    \caption{Variable importance for \subref{fig:CategorizedImportance01} 1, \subref{fig:CategorizedImportance02} 15, and \subref{fig:CategorizedImportance03} 30 days windowing}
+    \label{fig:CategorizedImportance}
+\end{figure}
+
 In general terms, RF had a better results of MAE, if compared to other models. SVR presented the worst result, while both Stacking Regressor and Decision Tree had similar results.
@@
 
-%\subsection{Machine learning models} \label{subsec:pre-processamento}
-
-%Parte a ser entregue até o dia 21/06.
-
-%\section{Discussion}
-
-%Discutir os resultados obtidos. Parte a ser entregue até o dia 21/06.
-
-%\section{Conclusion}
-
 \bibliographystyle{sbc}

### 2021-06-26 model_load.py
Gráfico de importâncias categorizado em escala logaritmica

@@ -146,9 +146,10 @@ def plot_save_categorized_importance(
 
         plt.plot(
             aaaa.loc[idx[:, windowing, 'mean']][column].index,
-            aaaa.loc[idx[:, windowing, 'mean']][column],
+            aaaa.loc[idx[:, windowing, 'mean']][column].clip(.001),
             label=column_label(column),
         )
+        plt.yscale('log')
 
     plt.ylim(lim)
     plt.ylabel(y_label)
@@ -169,7 +170,7 @@ height = 5
 
 # %%
 plot_save_categorized_importance(windowing=1,
-                                 lim=(-0.2, 2.6),
+                                 lim=(1E-3, 1E1),
                                  x_label='Shift (days)',
                                  y_label='Importance',
                                  width=width,
@@ -178,7 +179,7 @@ plot_save_categorized_importance(windowing=1,
 
 # %%
 plot_save_categorized_importance(windowing=15,
-                                 lim=(-0.2, 2.6),
+                                 lim=(1E-3, 1E1),
                                  x_label='Shift (days)',
                                  y_label='Importance',
                                  width=width,
@@ -187,7 +188,7 @@ plot_save_categorized_importance(windowing=15,
 
 # %%
 plot_save_categorized_importance(windowing=30,
-                                 lim=(-0.2, 2.6),
+                                 lim=(1E-3, 1E1),
                                  legend_loc='upper left',
                                  x_label='Shift (days)',
                                  y_label='Importance',

### 2021-06-26 model_load.py
Histogramas dos dados do reservatório

@@ -487,3 +487,19 @@ for shift in [1, 29]:  # range(1, 30, 7):
     plt.legend(loc='upper right')
     plt.savefig(f'paper/Graphs/Comparisson_d{shift:02d}.svg')
     plt.show()
+
+
+# %%
+plt.style.use('default')
+plt.gca().xaxis.set_major_locator(plt.MultipleLocator(500))
+plt.hist(regression_models[1]['windowed_data'][1]['streamflow'], bins=20, color='w', edgecolor='black')
+plt.axvline(np.median(regression_models[1]['windowed_data'][1]['streamflow'].values), color='b', linestyle='dashed')
+plt.xlabel('Streamflow ($m^3/s$)')
+# plt.xticks(ticks=500)
+# %%
+plt.style.use('default')
+plt.hist(regression_models[1]['windowed_data'][1]['level'], bins=20, color='w', edgecolor='black')
+plt.axvline(np.median(regression_models[1]['windowed_data'][1]['level'].values), color='b', linestyle='dashed')
+plt.xlabel('Streamflow ($m^3/s$)')
+
+# %%

### 2021-06-28 paper/paper.tex
Pequena correção em um parágrafo
Melhora no posicionamento dos caption dos gráficos de importância
Label nas seções

@@
 \section{Introduction}
+\label{sec:Introduction}
 
@@
 
-In Section 2, the related bibliography regarding machine learning models in the context of streamflow prediction is introduced.
-The study area and data used for the modeling are presented in Section 3, while the pre-processing data and results are described in Section 4, followed by a analysis about the results in Section 5 and conclusion in Section 6.
+In Section \ref{sec:RelatedBibliography}, the related bibliography regarding machine learning models in the context of streamflow prediction is introduced.
+The study area and data used for the modeling are presented in Section \ref{ssec:StudyAreaDecriptiveAnalysis}, while the pre-processing data and results are described in Section \ref{sec:DataPreprocessing}, followed by a analysis about the results in Section \ref{sec:ResultsDiscussion} and conclusion in Section \ref{sec:Conclusion}.
 
 \section{Related bibliography}
+\label{sec:RelatedBibliography}
 
@@
 \section{Material and methods}
+\label{sec:MateriaMethods}
 
@@
 \subsection{Study area and descriptive analysis}
+\label{ssec:StudyAreaDecriptiveAnalysis}
+
 We chose Três Marias reservoir for two main reasons.
@@
 \subsection{Data}
+\label{ssec:Data}
 
@@
 \subsection{Methods and evaluation criteria}
+\label{ssec:MethodsEvaluation}
 
@@
     \centering
-  %   \includegraphics[width=1\linewidth]{Diagrams/Pipelines.PNG}
-    {\small\includesvg[width=.8\linewidth]{Diagrams/gridsearch.drawio}}
+    {\footnotesize\includesvg[width=.8\linewidth]{Diagrams/gridsearch.drawio}}
     \caption{Pipeline setup for machine learning model search}
@@
 \section{Data preprocessing}
+\label{sec:DataPreprocessing}
 
@@
 \subsection{Data reduction and integration}
+\label{ssec:DataReductionIntegration}
 
@@
 
-\subsection{Data transformation} \label{sec:DataTransformation}
+\subsection{Data transformation}
+\label{sec:DataTransformation}
 
@@
 \section{Results and discussion}
+\label{sec:ResultsDiscussion}
 
@@
         \includesvg{Graphs/CategorizedImportance_w01}
-        \captionsetup{margin={.15\textwidth, 0cm}}
+        \captionsetup{margin={.17\textwidth, 0cm}}
         \caption{1 day windowing}
@@
         \includesvg{Graphs/CategorizedImportance_w15}
-        \captionsetup{margin={.15\textwidth, 0cm}}
+        \captionsetup{margin={.17\textwidth, 0cm}}
         \caption{15 days windowing}
@@
         \includesvg{Graphs/CategorizedImportance_w30}
-        \captionsetup{margin={.23\linewidth, 0cm}}
+        \captionsetup{margin={.26\linewidth, 0cm}}
         \captionsetup{justification=justified,singlelinecheck=false}
@@
 \section{Conclusion}
+\label{sec:Conclusion}
 

### 2021-06-28 model_load.py
Gráfico de comparação stacked

@@ -446,7 +446,7 @@ plot_save_score(windowing=30,
 
 # %%
 date_start = datetime.date(2000, 1, 1)
-date_end = datetime.date(2001, 1, 1)
+date_end = datetime.date(2002, 1, 1)
 interpolate_steps = 1
 time_slice = slice(date_start, date_end)
 time_slice_interpolate = slice(date_start, date_end, interpolate_steps)
@@ -487,6 +487,43 @@ for shift in [1, 29]:  # range(1, 30, 7):
     plt.legend(loc='upper right')
     plt.savefig(f'paper/Graphs/Comparisson_d{shift:02d}.svg')
     plt.show()
+
+# %%
+width = 15
+height = 8
+fig, ax = plt.subplots()
+plt.gcf().set_size_inches(cm_to_inches(width), cm_to_inches(height))
+ax.plot(original_data.loc[time_slice].index.values, interpolated_original_y, label='True')
+for shift in [1, 29]:  # range(1, 30, 7):
+    cv_predict = cross_val_predict(
+        regression_models[shift]['best_estimator'],
+        regression_models[shift]['best_windowed_data'],
+        regression_models[shift]['best_windowed_data']['streamflow'],
+        n_jobs=-1)
+
+    predicted_values = pandas.DataFrame(
+        cv_predict,
+        columns=['predict'],
+        index=regression_models[shift]['best_windowed_data'].index)
+
+    predict_X_y_spline = scipy.interpolate.make_interp_spline(
+        predicted_values.loc[time_slice_interpolate].index.values,
+        predicted_values.loc[time_slice_interpolate].values
+    )
+
+    interpolated_predicted_y = predict_X_y_spline(predicted_values.loc[time_slice].index).clip(min=0)
+    ax.plot(
+        predicted_values.loc[time_slice].index,
+        interpolated_predicted_y,
+        label=f'{shift}',
+        # linestyle='dashed',
+        linewidth=.85
+    )
+
+ax.legend(loc='best')
+fig.autofmt_xdate()
+plt.savefig(f'paper/Graphs/StackedComparisson.svg')
+plt.show()
 
 
 # %%

### 2021-06-28 model_load.py
Gráfico de comparação com valores para todo o período do dataset

@@ -10,7 +10,7 @@ import scipy.interpolate
 import scipy.stats
 from matplotlib.transforms import Affine2D
 from sklearn.model_selection import cross_val_predict
-from sqlalchemy.sql.expression import label
+from sqlalchemy.sql.expression import label, true
 
 from source.project_utils.constants import targets_models
 
@@ -522,7 +522,61 @@ for shift in [1, 29]:  # range(1, 30, 7):
 
 ax.legend(loc='best')
 fig.autofmt_xdate()
+plt.grid(True)
 plt.savefig(f'paper/Graphs/StackedComparisson.svg')
+plt.show()
+
+
+# %%
+date_start = datetime.date(1990, 1, 1)
+date_end = datetime.date(2010, 1, 1)
+interpolate_steps = 1
+time_slice = slice(date_start, date_end)
+time_slice_interpolate = slice(date_start, date_end, interpolate_steps)
+
+original_data = regression_models[1]['best_windowed_data']['streamflow']
+
+predict_X_y_spline = scipy.interpolate.make_interp_spline(
+    original_data.loc[time_slice_interpolate].index.values,
+    original_data.loc[time_slice_interpolate].values
+)
+interpolated_original_y = predict_X_y_spline(original_data.loc[time_slice].index).clip(min=0)
+# %%
+width = 15
+height = 8
+fig, ax = plt.subplots()
+plt.gcf().set_size_inches(cm_to_inches(width), cm_to_inches(height))
+ax.plot(original_data.loc[time_slice].index.values, interpolated_original_y, label='True')
+for shift in [1, 29]:  # range(1, 30, 7):
+    cv_predict = cross_val_predict(
+        regression_models[shift]['best_estimator'],
+        regression_models[shift]['best_windowed_data'],
+        regression_models[shift]['best_windowed_data']['streamflow'],
+        n_jobs=-1)
+
+    predicted_values = pandas.DataFrame(
+        cv_predict,
+        columns=['predict'],
+        index=regression_models[shift]['best_windowed_data'].index)
+
+    predict_X_y_spline = scipy.interpolate.make_interp_spline(
+        predicted_values.loc[time_slice_interpolate].index.values,
+        predicted_values.loc[time_slice_interpolate].values
+    )
+
+    interpolated_predicted_y = predict_X_y_spline(predicted_values.loc[time_slice].index).clip(min=0)
+    ax.plot(
[…]

### 2021-06-28 paper/paper.tex
Sobre o gráfico de comparação no paper

@@
 in which $n$ means the number of non-missing data, $i$ is the variable, $X_{pred}$ and $X_{i}$ mean the predicted and the actual value, respectively.
-Conversely, $R^2$ represents the proportion of the variance in the dependent variable which is explained by the regression model, given by the Equation
+Conversely, $R^2$ represents the proportion of the variance in the dependent variable which is explained by the regression model, and is given by the Equation
     \begin{equation}
@@
     \end{equation}
-where $X_{mean}$ is $\frac{1}{n}\sum_{i=0}^{n-1}X_{i}$ and $X_{pred}$ and $X_{i}$ are the predicted and it's corresponding true value, while $n$ is the size of the test sample.
+where $X_{mean}$ is $\frac{1}{n}\sum_{i=0}^{n-1}X_{i}$, and $X_{pred}$ and $X_{i}$ are the predicted and it's corresponding true value, while $n$ is the size of the test sample.
 R-square is a indicator of goodness, indicating how well the model can predict unseen samples.
@@
 
-For month variable we apply one-hot encoding, which is often used for categorical variables.
+For the month variable, we applied one-hot encoding, which is often used for categorical variables, when the regressor was sensitive to the scale of the attributes, such as SVR, or Ordinal encoding, otherwise.
 
@@
 Having tested horizon prediction of streamflow and water level varying from 1 to 29 days, and windowing at intervals of 1, 15 and 30 days for the climate variables, results indicated better fit for horizon of one day and window of 30 days.
-However, the models presented different results of R\textsuperscript{2} when they are compared to each other, as shown in Figures \ref{fig:r2_1}, \ref{fig:r2_15} and \ref{fig:r2_30}.
+However, the models presented different results of R\textsuperscript{2} when they are compared to each other, as shown in Figure~\ref{fig:r2}.
 
@@
 
+From an evaluation of Figure~\ref{fig:StackedComparisson}, which demonstrates the true versus predicted values for 1 and 29 days of prediction horizon, it is evident that the model trained for 1 day of horizon presents greater sensitivity to sudden changes in reservoir streamflow, while that the model with the longest horizon time, despite being able to follow the streamflow trend, does not faithfully represent the peaks present.
+Another point that can be justified by the presence of these peaks is the relatively large value found for the MAE since the model ends up being highly penalized by sudden changes in streamflow.
+On the other hand, an R-squared metric, which considers the amplitude of the actual data, ends up not being penalized in the same way.
+Figure~\ref{fig:StackedComparisson} presents data for a two-year window from 2000 to 2002, so data can be more easily visualized.
+For a visualization for the whole data range, see Appendix~\ref{sec:TrueVersusPredicted}
+
+\begin{figure}[htbp]
+    \centering
+    \includesvg{Graphs/StackedComparisson}
+    \caption{Comparison between true and predicted values for a prediction     horizon of 1 and 29 days. Visualization for a 2 year window, from 2000 to 2002}
+    \label{fig:StackedComparisson}
+\end{figure}
+
 After 8 days of prediction horizon, the R\textsuperscript{2} tends to be constant in 15 and 30 days-windowing cases.
@@
 The Stacking Regressor and decision tree produced similar values of MAE.
-Surprisingly, RF presented a slight increase of MAE for greater days of lagging, and for the case of 15 and 30 days-windowing, while the other models presented a trend of having a decline.
+Surprisingly, RF presented a slight increase of MAE for increased prediction horizon, and for the case of 15 and 30 days-windowing, while the other models presented a trend of having a decline.
 
@@
 That is important in solving water resources management problems and improve the streamflow prediction.
+
+As the prediction horizon increases, the tested models become less sensitive to drastic changes in the streamflow, not being able to predict peaks, but keeping the ability to describe the overall tendency.
 

### 2021-06-28 model_evaluate.py
Mudança no nome do arquivo de load para evaluate, para representar melhor o que o código faz

@@ -1,0 +1,596 @@
+# %%
+import datetime
+from typing import Tuple, Union
+
+import joblib
+import matplotlib.pyplot as plt
+import numpy as np
+import pandas
+import scipy.interpolate
+import scipy.stats
+from matplotlib.transforms import Affine2D
+from sklearn.model_selection import cross_val_predict
+from sqlalchemy.sql.expression import label, true
+
+from source.project_utils.constants import targets_models
+
+pandas.set_option('display.max_columns', 51)
+
+plt.style.use('default')
+
+idx = pandas.IndexSlice
+
+# %%
+# Carregar o modelo
+debug = False
+models_path = 'model/'
+if debug:
+    models_path += 'debug/'
+
+# %%
+regression_models = {}
+targets_models = range(1, 30, 7)
+for target in targets_models:
+    regression_models[target] = joblib.load(models_path + f'streamflow-s_d{target:02d}.pkl')
+
+# %%
+ImportancesDataFrame = joblib.load(models_path + 'importances.pkl')
+
+#
+# %%
+
+
+def cm_to_inches(cm):
+    return cm / 2.54
+
+
+def column_label(column: Tuple[str, Union[str, int]]):
+    if isinstance(column, str) == 1:
+        return column.replace('_', ' ').capitalize()
+
+    label = column[0].replace('_', ' ').capitalize()
+    if column[1] != "":
+        label += f' River {column[1]:02d}'
+
+    return label
+
+
+def plot_save_importance(
+        windowing=1,
+        lim=None,
+        legend_loc=0,
+        save_folder: str = '.',
+        x_label=None,
+        y_label=None,
+        width=6.4,
+        height=4.8):
+    if not save_folder.endswith('/'):
+        save_folder += '/'
+    aaaa = ImportancesDataFrame.copy().transpose()
+    aaaa[('mean', windowing, 'mean')] = [row.mean()
+                                         for row in ImportancesDataFrame.transpose().loc[idx[:, :, :], idx[:, windowing, 'mean']].values]
+    aaaa = aaaa.sort_values(('mean', windowing, 'mean'), ascending=False)[:5]
+    aaaa = aaaa.drop(columns=('mean', windowing, 'mean')).transpose()
+    for column in aaaa.loc[idx[:, windowing, 'mean']].columns:
+
+        plt.plot(
+            aaaa.loc[idx[:, windowing, 'mean']][column].index,
[…]

### 2021-06-28 paper/paper.tex
Adoção das correções feitas no overleaf

@@
 Hydropower operation and planning require streamflow forecasts.
-Machine learning methods combined with atmospheric climate variables complement physical models, particularly, where in-situ observations are not available.
+Machine learning (ML) methods combined with atmospheric climate variables complement physical models, particularly, where in-situ observations are not available.
 To make predictions of streamflow of a reservoir, we test four ML techniques: Support Vector Regressor (SVR), Simple Decision Tree, Random Forest (RF), and Stacking Regressor, which combines SVR, RF, and Ridge.
-Also, we add prediction horizon varying from 0 to 29 days for streamflow, and windowing of 1, 15, and 30 days for atmospheric climate variables.
+Also, we add prediction horizon varying from 1 to 29 days for streamflow, and windowing of 1, 15, and 30 days for atmospheric climate variables.
 Results indicate a better fit in the RF model.
@@
 
-In this work, we make short-term forecast streamflow using Três Marias reservoir as study case.
+In this work, we explore machine learning methods for short-term forecast streamflow using Três Marias reservoir as a case study.
+The objective here is to evaluate the performance of Support Vector Regressor (SVR), Simple Decision Tree, Random Forest (RF), and Stacking Regressor, which combines SVR, RF, and Ridge, in different prediction horizons and climate variable windowing.
 We use atmospheric variables as precipitation, evaporation, temperature, and surface runoff, available in Projeta project \cite{inpe}, and observed data of streamflow and water level of Três Marias reservoir \cite{onsnivel,onsvazao}.
-We compared four machine learning methods in order to access the best performance of short-term predictions.
-Moreover, we evaluate the impact of adding a horizon prediction of streamflow, as well as, windowing the climate variables.
 
@@
 Stacking ensemble ML algorithm learns how to combine the predictors from other well-performing models.
-In our case, we combined both SVR, RF models and Ridge regressors.
+In our case, we combine both SVR, RF models and Ridge regressors.
 
@@
 Historical data of streamflow itself as input variable was present in \cite{li}.
-In order to approach the natural structure of a reservoir, we included in our work the variables runoff and evaporation, besides precipitation, evaporation, streamflow and water level.
+In order to approach the natural structure of a reservoir, we include in our work the variables runoff and evaporation, besides precipitation, evaporation, streamflow and water level.
 
@@
 We evaluated the performance of each model adding a windowing for climate variables in intervals of 1, 15, and 30 days.
-Furthermore, we tested horizon predictions varying from 0 to 29 days.
+Furthermore, we tested horizon predictions varying from 1 to 29 days.
 
@@
 We used a combination of parameters to tune the model with an R-squared score as a metric, leveraging the GridSearchCV scikit-learn tool.
-In GridSearchCV, we define some constraints of each parameter, available in Appendix B.
+In GridSearchCV, we define some constraints of each parameter, detailed in Appendix B.
 We adopted a cross-validation strategy to optimize these parameters and minimize the risk of overfitting.
 This was made to avoid testing regressor parameters on the same data it was trained for.
-The adopted approach was k-fold cross-validation, where the data set was divided into k groups - $k$ was set to 10 in this study - the regressors were trained in $k-1$ of these groups, and the fold left was used to test.
-As a consequence, the test was made using data the regressor have never seen while training.
+The adopted approach was K-Fold cross-validation, where the data set was divided into k groups - $k$ was set to 10 in this study - the regressors were trained in $k-1$ of these groups, and the fold left was used to test.
+As a consequence, the test data used by the regressors were new for every new training session.
 We calculate the final score as the average of the scores obtained for each test group.
@@
 
-All variables, except month variable, were normalized when evaluating regressors that are sensitive to data scale, in order to prevent them from being dominated by variables with higher order of magnitude.
+All variables, except month, were normalized when evaluating regressors that are sensitive to data scale, in order to prevent them from being dominated by variables with higher order of magnitude.
 We used the Min-Max normalization given by the Equation
@@
 Having tested horizon prediction of streamflow and water level varying from 1 to 29 days, and windowing at intervals of 1, 15 and 30 days for the climate variables, results indicated better fit for horizon of one day and window of 30 days.
-However, the models presented different results of R\textsuperscript{2} when they are compared to each other, as shown in Figure~\ref{fig:r2}.
+However, the models presented different results of R-squared when they are compared to each other, as shown in Figure~\ref{fig:r2}.
 
@@
     \end{subfigure}
-    \caption{R\textsuperscript{2} for \subref{fig:r2_1} 1, \subref{fig:r2_15} 15, and \subref{fig:r2_30} 30 days windowing}
+    \caption{R-squared for \subref{fig:r2_1} 1, \subref{fig:r2_15} 15, and \subref{fig:r2_30} 30 days windowing}
     \label{fig:r2}
@@
 
-According to the increase of prediction horizon of streamflow up to 8 days, the R\textsuperscript{2} had an accentuated decrease in all models in 1 day-windowing.
-This trend is kept in 15 and 30 days windowing only for Decision Tree, while for SVR and Stacking Regressor the decrease is smaller and for Random Forest, R\textsuperscript{2} presented a growth.
+According to the increase of prediction horizon of streamflow up to 8 days, the R-squared had an accentuated decrease in all models in 1 day-windowing.
+This trend is kept in 15 and 30 days windowing only for Decision Tree, while for SVR and Stacking Regressor the decrease is smaller and for Random Forest, R-squared presented a growth.
 
@@
 On the other hand, an R-squared metric, which considers the amplitude of the actual data, ends up not being penalized in the same way.
-Figure~\ref{fig:StackedComparisson} presents data for a two-year window from 2000 to 2002, so data can be more easily visualized.
+Figure~\ref{fig:StackedComparisson} presents data for a two-year window from 2000 to 2002, so it can be more easily visualized.
 For a visualization for the whole data range, see Appendix~\ref{sec:TrueVersusPredicted}
@@
 
-After 8 days of prediction horizon, the R\textsuperscript{2} tends to be constant in 15 and 30 days-windowing cases.
+After 8 days of prediction horizon, the R-squared tends to be constant in 15 and 30 days-windowing cases.
 In 1 day-windowing case, the models presented an unexpected performance, since the models had a turning point in 15 days-horizon, in which the performance reached its lowest point and rose again.
 
-Random forest outperforms in all considered cases of prediction horizon and windowing in terms of R\textsuperscript{2}.
-Decision Tree Regressor has the worst fit, while SVR and Stacking Regressor had similar results, although Stacking Regressor slightly presents higher values of R\textsuperscript{2}.
+Random forest outperforms in all considered cases of prediction horizon and windowing in terms of R-squared.
+Decision Tree Regressor has the worst fit, while SVR and Stacking Regressor had similar results, although Stacking Regressor slightly presents higher values of R-squared.
 
@@
 
-In general terms, RF had a better results of MAE, if compared to other models. SVR presented the worst result, while both Stacking Regressor and Decision Tree had similar results.
+In general terms, RF had a better results of MAE and R-squared, if compared to other models. SVR presented the worst result, while both Stacking Regressor and Decision Tree had similar results.
 
@@
 
-In this work, we tested four methods to forecast streamflow using Três Marias reservoir as a study case.
-Results indicated that RF had the best fit over the models.
-For one-day prediction horizon for streamflow and water level , models presented R-squared values in the range 0.7 to 0.9, but that decreases according to the increase of  prediction horizon.
+In this work, we tested four methods to forecast streamflow using Três Marias reservoir as a case study.
+Empirical results obtained from these models showed the performance not only over the models, but also changing the prediction horizon of streamflow, and the window period of climate variables.
+That is important in solving water resources management problems and improve streamflow prediction.
+
+Results indicated that RF had the best fit over the other models.
+For one-day prediction horizon for streamflow, models presented R-squared values in the range 0.7 to 0.9, but that decreases according to the increase of  prediction horizon.
 With regards to windowing of climate variables, the best fits occur when considering a 30 days-windowing, which might be related to the delay of the physical process impact of all climate variables in the entire sub-basin on the reservoir upstream.
@@
 SVR had the highest values of MAE, while RF had the lowest values.
-The Stacking Regressor and decision tree produced similar values of MAE.
+The Stacking Regressor and decision tree produced the 2nd and 3rd best MAE results, respectively.
 Surprisingly, RF presented a slight increase of MAE for increased prediction horizon, and for the case of 15 and 30 days-windowing, while the other models presented a trend of having a decline.
-
-Empirical results obtained from these models showed the performance not only over the models, but also changing the lag of streamflow and water level, and the window period of climate variables.
-That is important in solving water resources management problems and improve the streamflow prediction.
 

### 2021-06-28 model_evaluate.py
Mudança no tamanho dos gráficos

@@ -128,6 +128,18 @@ plot_save_importance(windowing=30,
 
 
 # %%
+col_symbol = {
+    'Temperature': 'T',
+    'Precipitation': 'P',
+    'Evaporation': 'E',
+    'Surface runoff': 'R',
+    'Shifted streamflow': 'SF',
+    'Shifted level': 'WL',
+    'Month': 'M'
+}
+# %%
+
+
 def plot_save_categorized_importance(
         windowing=1,
         lim=None,
@@ -147,7 +159,7 @@ def plot_save_categorized_importance(
         plt.plot(
             aaaa.loc[idx[:, windowing, 'mean']][column].index,
             aaaa.loc[idx[:, windowing, 'mean']][column].clip(.001),
-            label=column_label(column),
+            label=col_symbol[column_label(column)],
         )
         plt.yscale('log')
 
@@ -157,7 +169,7 @@ def plot_save_categorized_importance(
     plt.xticks(ImportancesDataFrame.loc[idx[:, windowing, 'mean']][column].index)
 
     if legend_loc is not None:
-        plt.legend(loc=legend_loc, bbox_to_anchor=(1.05, 1))
+        plt.legend(loc=legend_loc, bbox_to_anchor=(1.05, 1), ncol=2)
     plt.gcf().set_size_inches(cm_to_inches(width), cm_to_inches(height))
     plt.grid(True)
     plt.tight_layout()
@@ -166,7 +178,7 @@ def plot_save_categorized_importance(
 
 # %%
 width = 7.3
-height = 5
+height = 4
 
 # %%
 plot_save_categorized_importance(windowing=1,
@@ -189,10 +201,10 @@ plot_save_categorized_importance(windowing=15,
 # %%
 plot_save_categorized_importance(windowing=30,
                                  lim=(1E-3, 1E1),
-                                 legend_loc='upper left',
+                                 legend_loc='best',
                                  x_label='Shift (days)',
                                  y_label='Importance',
-                                 width=width + 5.2,
+                                 width=width + 4.75,
                                  height=height,
                                  save_folder='paper/graphs')
 
@@ -317,7 +329,7 @@ def plot_save_score(
 
 # %%
 width = 7.3
-height = 4.5
+height = 4
 # %%
 plot_save_score(windowing=1,
                 scorer='R2',
[…]

### 2021-06-28 paper/paper.tex
Figuras agora posicionadas apenas como htb
boxplot para o appendix, retirada do parágrafo sobre

@@
 
-\begin{figure}[htbp]
+\begin{figure}[htb]
   \centering
-  \includegraphics[width=0.9\linewidth]{Figures/mapa.pdf}
+  \includegraphics[width=0.6\linewidth]{Figures/mapa.png}
+  \includesvg{Figures/mapa_legend.drawio}
   \caption{Alto São Francisco sub-basin.}
@@
 All variables are shown in Table~\ref{tab:predictors}.
-For further information of data exploration, see Appendix A.
+For further information of data exploration, see Appendix~\ref{sec:DataExploration}.
 
-\begin{table}[htbp]
+\begin{table}[htb]
 \centering
@@
 
-\begin{figure}[htbp]
+\begin{figure}[htb]
     \centering
@@
 Then, we transformed that to their original scale.
-
-While only the scale in the values changed with the normalization procedure, it is easier to identify the variation in mean and standard deviation between variables, as shown in Figure \ref{fig:Normalized} for the inputs in the first river.
-
-\begin{figure}[htbp]
-  \centering
-  \includegraphics[width=0.8\linewidth, trim=0cm 0 0 .7cm,clip=true]{Figures/Normalização.png}
-  \caption{Boxplot of inputs from normalized data (left) and raw data (rifht)}
-  \label{fig:Normalized}
-\end{figure}
-
 For the month variable, we applied one-hot encoding, which is often used for categorical variables, when the regressor was sensitive to the scale of the attributes, such as SVR, or Ordinal encoding, otherwise.

### 2021-06-28 model_load.py

@@ -1,596 +1,0 @@
-# %%
-import datetime
-from typing import Tuple, Union
-
-import joblib
-import matplotlib.pyplot as plt
-import numpy as np
-import pandas
-import scipy.interpolate
-import scipy.stats
-from matplotlib.transforms import Affine2D
-from sklearn.model_selection import cross_val_predict
-from sqlalchemy.sql.expression import label, true
-
-from source.project_utils.constants import targets_models
-
-pandas.set_option('display.max_columns', 51)
-
-plt.style.use('default')
-
-idx = pandas.IndexSlice
-
-# %%
-# Carregar o modelo
-debug = False
-models_path = 'model/'
-if debug:
-    models_path += 'debug/'
-
-# %%
-regression_models = {}
-targets_models = range(1, 30, 7)
-for target in targets_models:
-    regression_models[target] = joblib.load(models_path + f'streamflow-s_d{target:02d}.pkl')
-
-# %%
-ImportancesDataFrame = joblib.load(models_path + 'importances.pkl')
-
-#
-# %%
-
-
-def cm_to_inches(cm):
-    return cm / 2.54
-
-
-def column_label(column: Tuple[str, Union[str, int]]):
-    if isinstance(column, str) == 1:
-        return column.replace('_', ' ').capitalize()
-
-    label = column[0].replace('_', ' ').capitalize()
-    if column[1] != "":
-        label += f' River {column[1]:02d}'
-
-    return label
-
-
-def plot_save_importance(
-        windowing=1,
-        lim=None,
-        legend_loc=0,
-        save_folder: str = '.',
-        x_label=None,
-        y_label=None,
-        width=6.4,
-        height=4.8):
-    if not save_folder.endswith('/'):
-        save_folder += '/'
-    aaaa = ImportancesDataFrame.copy().transpose()
-    aaaa[('mean', windowing, 'mean')] = [row.mean()
-                                         for row in ImportancesDataFrame.transpose().loc[idx[:, :, :], idx[:, windowing, 'mean']].values]
-    aaaa = aaaa.sort_values(('mean', windowing, 'mean'), ascending=False)[:5]
-    aaaa = aaaa.drop(columns=('mean', windowing, 'mean')).transpose()
-    for column in aaaa.loc[idx[:, windowing, 'mean']].columns:
-
-        plt.plot(
-            aaaa.loc[idx[:, windowing, 'mean']][column].index,
[…]
